### 2020-06-14 analysis/002_cr_els_matching.Rmd
prepare parsing for invoice data

---
title: "Obtain metadata from Crossref Dump"
output: github_document
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

### Connect to database

```{r}
# connect to google bg where we imported the json lines Unpaywall dump
library(DBI)
library(bigrquery)
con <- dbConnect(
  bigrquery::bigquery(),
  project = "api-project-764811344545",
  dataset = "cr_dump_march_20"
)
```

### Journals associated with Elsevier

Determined by the Crossref member id (78).

```{sql connection=con, output.var="els_2020_jns"}
SELECT DISTINCT(ISSN)
FROM `api-project-764811344545.cr_dump_march_20.cr_08_20_expanded` 
WHERE member = '78' AND issued_year BETWEEN 2015 AND 2019
```

### Comparision with Crossref Dump

Journals from hybrid Elsevier list not indexed in Crossref

```{r}
library(tidyverse)
library(here)
els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
els_jns_df %>% 
  filter(!issn %in% els_2020_jns$ISSN, oa_type == "Hybrid")
```

Possible reasons:

a) Journal transfer including metadata ownership. Example: Canadian Association of Radiologists Journal (transferred from Elsevier to SAGE Publications as of 2020)

```{sql connection=con}
SELECT member, publisher, COUNT(DISTINCT(doi)) as articles
FROM `api-project-764811344545.cr_dump_march_20.cr_08_20_expanded` 
WHERE container_title = 'Canadian Association of Radiologists Journal'
GROUP BY member, publisher
```

However, there are cases, where metadata ownership was not fully transferred. Example: Acta Mathematica Scientia (transferred to Springer Nature as of 2019)

```{sql connection=con}
SELECT member, publisher, COUNT(DISTINCT(doi)) as articles
FROM `api-project-764811344545.cr_dump_march_20.cr_08_20_expanded` 
WHERE container_title = 'Acta Mathematica Scientia'
GROUP BY member, publisher
```

b) Journal started in 2020

### Create Crossref Elsevier subset in BQ

Upload list of Elsevier hybrid journals to BQ.

```{r}
els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
els_jn_hybrid <- els_jns_df %>% 
  filter(oa_type == "Hybrid")
if(!bq_table_exists("api-project-764811344545.cr_dump_march_20.elshybrid")) {
bq_els <- bq_table_create(
  "api-project-764811344545.cr_dump_march_20.elshybrid",
  fields =  els_jn_hybrid,
  friendly_name = "Elsevier hybrid journals May 2020",
  description = "The data was extracted from Elsevier APC pricing list, 
    downloaded 30 May 2020"
)
bq_table_upload(bq_els, els_jn_hybrid)
} else {
  NULL
}
```

Create table with Crossref metadata for these journals since 2008.

```{sql, connection=con}
CREATE OR REPLACE TABLE `api-project-764811344545.cr_dump_march_20.els_hybrid_cr` AS
SELECT * FROM `api-project-764811344545.cr_dump_march_20.cr_08_20_expanded` as cr
WHERE EXISTS ( 
  SELECT 1 FROM `api-project-764811344545.cr_dump_march_20.elshybrid` as els_hybrid 
  WHERE (cr.issn = els_hybrid.issn)
  )
```

Descriptive:

By members

```{sql, connection=con}
SELECT  member, publisher, COUNT(DISTINCT(doi)) as articles
FROM `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`
WHERE issued_year > 2014 and issued_year < 2020
GROUP BY member, publisher
ORDER BY articles DESC
```

License prevalence

```{sql, connection=con}
SELECT REGEXP_REPLACE(license.URL, "http://|https://", "") as license_url, COUNT(DISTINCT(doi)) as articles
FROM `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`, unnest(license) as license
WHERE issued_year > 2014 and issued_year < 2020 AND license.content_version = 'vor'
GROUP BY license_url
ORDER BY articles DESC
```

License prevalence by days with delay

```{sql, connection=con, output.var="els_license_delay"}
SELECT 
  REGEXP_REPLACE(license.URL, "http://|https://", "") as license_url,
  license.delay_in_days,
  COUNT(DISTINCT(doi)) as articles
FROM `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`, unnest(license) as license
WHERE issued_year > 2014 and issued_year < 2020 AND license.content_version = 'vor'
GROUP BY license_url, license.delay_in_days
ORDER BY articles DESC
```

Immediate Licenses

```{r}
els_license_delay %>%
  filter(delay_in_days == 0)
```

A few records have a date formats without day, which is used for the delayed calculation. We should therefore allow for a delay of up to  30 days.

```{r}
els_license_delay %>%
  filter(delay_in_days < 31) 
```

### Download open access articles in hybrid journals

```{sql, connection=con, output.var="els_hybrid_articles"}
SELECT *
FROM `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`, unnest(license) as license
WHERE (REGEXP_CONTAINS(license.URL, 'creativecommons') AND license.content_version = 'vor') AND 
  (issued_year BETWEEN 2015 and 2019) AND
  license.delay_in_days < 31
```

Backup

```{r}
require(jsonlite)
jsonlite::stream_out(els_hybrid_articles, file(here::here("data", "els_hybrid_articles.json")))
```