### 2020-06-14 analysis/003_cr_xml_fetch.Rmd

---
title: "003_ftx_mining.Rmd"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Obtain Elsevier ftxt links for OA articles

Libraries

```{r}
require(jsonlite)
require(tidyverse)
require(crminer)
require(xml2)
require(here)
```

Load Elsevier articles:

```{r}
els_hybrid <- jsonlite::stream_in(file(here::here("data", "els_hybrid_articles.json")))
```

get tdm links

```{r}
elsevier_tdm <- els_hybrid %>%
  select(doi, link) %>%
  unnest(link) %>%
  filter(content_type == "text/xml",
         intended_application == "text-mining",
         content_version == "vor")
```

Apply function on every XML full-text link

```{r}
tt <- purrr::map(sample(elsevier_tdm$URL, 10), .f = purrr::safely(elsevier_parse))
```