### 2020-06-14 analysis/003_cr_xml_fetch.Rmd
no delay info from crossref. instead, the open archive info from elsevier is used

---
title: "003_ftx_mining.Rmd"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Obtain Elsevier ftxt links for OA articles

Libraries

```{r}
require(tidyverse)
require(crminer)
require(xml2)
require(here)
source(here::here("R", "parse_els_xml.R"))
```

Load Elsevier articles:

```{r}
els_hybrid <- readr::read_csv(here::here("data", "els_tmd_links.csv"))
```

```{r}
tt <- plyr::llply(sample(elsevier_tdm$URL, 10), .f = purrr::safely(elsevier_parse), .progress = "text")
```