### 2020-07-25 analysis/004_EDA.Rmd
starting with summary table, formating issue.

---
title: "EDA Hybrid OA"
author: "Najko Jahn"
date: "6/15/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(tidyverse)
require(here)
els_invoicing <- readr::read_csv(here::here("data", "invoice_df.csv"))
els_tdm <- readr::read_csv(here::here("data", "els_tmd_links.csv"))
els_df <- els_tdm %>%
  select(doi, URL) %>%
  inner_join(els_invoicing, by = c("URL" = "tdm_url")) %>% 
  distinct()
```

### By sponsor type and archive

```{r}
els_df %>%
  count(oa_sponsor_type, oa_archive)
```

### By invoice recipient

```{r}
els_df %>%
  count(oa_sponsor_name) %>%
  arrange(desc(n))
```

### compare with existing table

```{r}
library(googlesheets4)
library(writexl)
googlesheets4::gs4_deauth()
gs_df <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1AqjCLn6VNGJpQ6zdmFSwuhFsDBIITSxijA0r0WLJZTQ/")
miss_funder <- els_df %>%
  filter(!oa_sponsor_name %in% gs_df$oa_sponsor_name) %>%
  group_by(oa_sponsor_name) %>%
  summarise(n = n_distinct(doi))

writexl::write_xlsx(miss_funder, here::here("data", "new_funders.xlsx"))
```

### Master dataset

```{r}
els_invoicing <- readr::read_csv(here::here("data", "invoice_df.csv"))
els_tdm <- readr::read_csv(here::here("data", "els_tmd_links.csv"))
hybrid_md <- readr::read_csv(here::here("data", "hybrid_md.csv"))
els_df <- els_tdm %>%
  select(doi, URL) %>%
  inner_join(els_invoicing, by = c("URL" = "tdm_url")) %>% 
  distinct()
els_md <- els_df %>% 
  rename(tdm_url = URL) %>% 
  inner_join(hybrid_md, by = "doi") %>% 
  distinct(doi, .keep_all = TRUE)
readr::write_csv(els_md, here::here("data", "els_analyses_main.csv"))
````