### 2020-10-25 analysis/009_pricing_exploration.Rmd
apc / funding type

---
title: "pricing differences"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
```

read invoicing data

```{r}
hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
```

apc price list

```{r}
els_jn <- readr::read_csv(here::here("data", "els_historic_jns.csv"))
```

APC development hybrid oa portflio

```{r}
els_jn %>%
  filter(oa_model == "Hybrid") %>%
  ggplot(aes(year, apc, group = year)) +
  geom_boxplot()
```

By invoicing

```{r}
my_df <- left_join(hybrid_df, els_jn, by = c("issn", "issued_year" = "year")) %>%
    mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  filter(!fct_source %in% c("Other", NA))

```

```{r}
ggplot(my_df, aes(factor(issued_year), apc,  color = fct_source)) +
  geom_boxplot()
```

All

```{r}
ggplot(my_df, aes(factor(fct_source), apc)) +
  geom_boxplot()
```