### 2021-01-27 analysis/manuscript.Rmd

---
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: apa6.sty
  html_document:
    fig_caption: yes
    force_captions: yes
    number_sections: no
    theme: readable
csl: apa.csl
bibliography: references.bib
---

```{r knitr_options, , echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300,
  fig.pos = "H", 
  out.extra = ""
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(formatC(x, big.mark = ",", format = "f", digits = 1, drop0trailing = TRUE))
    } else{
      return(x)
    }
  }
)
```

```{r libraries}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(gt)
library(janitor)
library(kableExtra)
library(stringr)
```

# Results{.unnumbered}

This section first presents the results of our analysis of hybrid OA uptake in Elsevier's journal portfolio with a view to licensing, disciplinary differences, and citation impact. Then, we present a descriptive analysis of Elsevier's invoicing data, highlighting licensing and disciplinary differences for different invoicing channels, institutional sectors, and individual invoice recipients.

```{r data}
# journal-level data
els_jns <- readr::read_csv(here::here("data", "els_historic_jns.csv"))
els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
# article-level data
hybrid_oa_volume <- readr::read_csv(here::here("data", "hybrid_oa_volume.csv"))
# invoicing analysis
curated_invoicing_df <- readr::read_csv(here::here("data", "curated_invoicing_data.csv"))
# open apc data
o_apc <- readr::read_csv(here::here("data", "o_apc.csv"))
```

## Uptake of Hybrid Open Access{.unnumbered}

### Overview{.unnumbered}

```{r}
# journals by year and oa model
els_jns_count <- els_jns %>%
  group_by(year, oa_model) %>%
  count(year) %>%
  group_by(year) %>%
  mutate(prop = n / sum(n)) %>%
  mutate(oa_model_fct = fct_rev(as_factor(oa_model)))
# overall count
 els_overall <- els_jns %>% 
  group_by(oa_model) %>% 
  summarise(n = n_distinct(issn))
```

```{r, rev_flipped}
flipped_jns <- els_jns %>%
  distinct(issn, oa_model) %>%
  group_by(issn) %>%
  filter(n() > 1)
flipped_jns_ind <- els_jns %>%
  filter(issn %in% flipped_jns$issn) %>%
  group_by(issn) %>%
  slice(which.max(year)) %>%
  ungroup() %>%
  count(oa_model)
```
```{r hybrid_oa_per_year}
# journal and article metrics **per year**
hybrid_oa_per_year <- hybrid_oa_volume %>%
  group_by(issued_year) %>%
  # distinct jns with at least one article per year
  filter(oa > 0) %>%
  summarise(
      # distinct jns with at least one article per year
    n_journals = n(),
    # article metrics
    articles_n = sum(articles),
    articles_mean = mean(articles),
    articles_sd = sd(articles),
    # oa articles metrics
    oa_n = sum(oa),
    oa_mean = mean(oa),
    oa_sd = sd(oa)
    ) %>%
  mutate(oa_uptake = oa_n / articles_n)
# oa uptake per year
oa_uptake_per_year <- hybrid_oa_volume %>% 
  filter(oa > 0) %>%
  mutate(prop = oa / articles) %>%
  group_by(issued_year) %>%
  summarise(oa_uptake_mean = mean(prop),
            oa_uptake_sd = sd(prop))
# merge indicators
ind_per_year <- inner_join(hybrid_oa_per_year, oa_uptake_per_year, by = "issued_year")

# five-years article volume
all_jn_articles <- hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    articles_p_jn = sum(articles)
    ) %>%
  ungroup() %>%
  summarise(
    n_journals = n(),
    articles_n = sum(articles_p_jn),
    articles_mean = mean(articles_p_jn),
    articles_sd = sd(articles_p_jn)
  )
# five-years OA article volume
all_oa_articles <- 
  hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    oa_p_jn = sum(oa)
    ) %>%
  ungroup() %>%
  summarise(
    oa_n = sum(oa_p_jn),
    oa_mean = mean(oa_p_jn),
    oa_sd = sd(oa_p_jn)
  )
# five-years OA uptake rate
oa_uptake <- 
  hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    articles_p_jn = sum(articles),
    oa_p_jn = sum(oa)
    ) %>%
  mutate(prop = oa_p_jn / articles_p_jn) %>%
  ungroup() %>%
  summarise(oa_uptake = sum(oa_p_jn) / sum(articles_p_jn),
            oa_uptake_mean = mean(prop),
            oa_uptake_sd = sd(prop)
  )
# bringing it all together
five_years_ind <- bind_cols(issued_year = "2015-19", all_jn_articles, all_oa_articles, oa_uptake)

ind_summary_table <- ind_per_year %>%
  mutate(issued_year = as.character(issued_year)) %>% 
  bind_rows(five_years_ind) %>%
  mutate(across(contains("uptake")) * 100) %>%
  pivot_longer(n_journals:oa_uptake_sd, names_to = "indicator", values_to = "count") %>%
  pivot_wider(names_from = issued_year, values_from = count)
```

Between 2015 and 2019, `r hybrid_oa_volume %>% distinct(issn) %>% nrow()` of `r els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` (`r  hybrid_oa_volume %>% distinct(issn) %>% nrow() / (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n))*100`%) hybrid journals published at least one OA article. Together these journals published `r five_years_ind %>% pull(oa_n)` OA articles, which represents `r five_years_ind %>% pull(oa_uptake)*100`% of their total article volume (n=`r five_years_ind %>% pull(articles_n)`). Table \ref{tab:overview} presents these findings for each year and the aggregated five-year period, illustrating a moderate growth in hybrid OA over time. With each year, the number of hybrid journals with at least one OA article increased, as did the number of immediate OA articles in these journals, which nearly doubled from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_n)` in 2015 to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_n)` in 2019. However, since the total number of articles published in these journals also grew over time, the relative share of OA articles in hybrid journals only increased slightly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake)*100`% to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake)*100`%.

```{r overview}
table_descriptives_ind <- tribble(
  ~ind_group, ~ind, ~` `,
    "Elsevier Hybrid Journals with ≥ 1 OA article", "n_journals", "Total",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_n", "Total ",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_mean", "Avg.",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_sd", "SD",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_n", "Total",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_mean", "Avg.",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_sd", "SD",
  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake", "Total (%)",
  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake_mean", "Avg.",
  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake_sd", "SD"
)
inner_join(table_descriptives_ind, ind_summary_table, by = c("ind" = "indicator")) %>%
  dplyr::select(-ind, -ind_group) %>%
  kableExtra::kbl(booktabs = TRUE, format.args = list(big.mark = ","), digits = 1, escape = TRUE,
                  caption = "Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019.",
                  position = "H"
) %>%
  kable_styling() %>%
  pack_rows("Elsevier Hybrid Journals with ≥ 1 OA article", 1, 1) %>%
  pack_rows("Articles in Elsevier Hybrid Journals with ≥ 1 OA Article", 2, 4) %>%
  pack_rows("OA Articles in Elsevier Hybrid Journals with ≥ 1 OA Article", 5, 7) %>%
  pack_rows("OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA Article", 8, 10)
```

```{r}
perc_df <- hybrid_oa_volume %>%
  mutate(prop = oa / articles) %>%
  filter(oa > 0) %>%
  group_by(issued_year) %>% 
  summarize(median = median(prop),
            p_10 = quantile(prop, 0.1),
            p_95 = quantile(prop, 0.95),
            iqr = quantile(prop, 3/4) - quantile(prop, 1/4)) %>%
  ungroup()
```

Figure \ref{fig:boxuptake} presents the hybrid OA uptake by year, showing considerable variation among Elsevier's hybrid journals. The box-and-whiskers plot visualizes the median hybrid OA uptake (horizontal line inside the box), the range of the majority of journals (the box), and individual outliers (dots). Notably, the upper quartiles and whiskers stretch farther over the years, which indicates that the range of uptake among journals with an above-average proportion of OA articles increased over the years. Nevertheless, the prevalence of hybrid OA remained relatively low; in 2019, 95% of Elsevier hybrid journals published ≤`r perc_df %>% filter(issued_year == 2019) %>% pull(p_95) * 100`% of their articles OA.

```{r boxuptake, fig.cap = "Open access uptake per Elsevier hybrid journal in percent by year, visualized as box-and-whiskers plot. The Y-axis is limited to an open access share of 20\\%."}
hybrid_oa_volume %>%
  mutate(prop = oa / articles) %>%
  filter(oa > 0) %>%
ggplot(aes(factor(issued_year), prop)) + 
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) + 
  coord_cartesian(ylim=c(0,0.2)) +
  theme_minimal_hgrid() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L),
      expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Publication Year",
       y = "Hybrid OA Uptake")
```

### Share of the Total Volume published in OA{.unnumbered}

```{r oa_volume}
# see analysis/006_license.Rmd how the data was obtained

license_df <- readr::read_csv(here::here("data", "year_per_oa_type_and_license.csv"))
# all oa
all_oa <- license_df %>% 
  group_by(type) %>% 
  summarise(n = sum(articles)) %>%
  pull(n) %>%
  sum()
# all articles
all_els <- license_df %>%
  distinct(all_articles, issued_year) %>%
  pull(all_articles) %>%
  sum()
# summary by type
summary_oa_type <- 
  license_df %>% 
  group_by(type) %>% 
  summarise(n = sum(articles)) %>%
  mutate(prop = n / sum(n))
# mirror journals
mirror_jn <- readr::read_csv(here::here("data", "x_journals_volume.csv"))
```

During the same period, Elsevier published `r all_oa` OA articles in total (i.e., in hybrid and full OA journals), representing `r all_oa / all_els * 100`% of the total article volume (n=`r all_els`). Looking at Figure \ref{fig:oa_volume_fig}, it is apparent that the OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged behind that of Elsevier's Open Archive program (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) and full OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`, `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r sum(mirror_jn$all_articles)` articles published in `r length(unique(mirror_jn$container_title))` mirror journals.

```{r oa_volume_fig, fig.cap="Open Access (OA) in Elsevier journals in terms of article volume between 2015-2019. Figure A presents the aggregated OA article volume of Elsevier journals from 2015-2019 by OA type. Figure B presents the share of OA articles relative to Elsevier's overall article output by OA type and open content license from 2015-2019."}
# see analysis/006_license.Rmd how the figure was created

knitr::include_graphics(here::here("figure", "license_portfolio.png"))
```

### License Prevalence{.unnumbered}

```{r cc_by_by_type}
cc_by_by_type <- license_df %>%
  group_by(type, license_code) %>%
  summarise(n = sum(articles)) %>%
  mutate(prop = n / sum(n)) %>%
  filter(license_code == "CC BY")
```

As far as we observed, OA articles in Elsevier hybrid journals were published under two possible CC licenses--CC BY or CC BY-NC-ND. CC BY allows others to distribute, remix, adapt, and build upon the licensed work, including for commercial purposes, as long as the original author is credited. In contrast, CC BY-NC-ND is less permissive as it prohibits commercial reuse and derivatives. In the five-year period from 2015-2019, the proportion of hybrid OA articles published under a CC BY-NC-ND license marginally increased and maintained a greater share than that of CC BY (see Figure \ref{fig:oa_volume_fig}-B). Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid journals had the highest number and proportion of openly available articles licensed under CC BY (n=`r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(prop) *100`%) compared to full OA journals  (n=`r filter(cc_by_by_type, type == "full_oa") %>% pull(n)`, `r filter(cc_by_by_type, type == "full_oa") %>% pull(prop) *100`%) and Elsevier's delayed OA program (n=`r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(prop) *100`%) . Most  delayed OA articles were provided under an Elsevier user license (Els-User), which prohibits reuse for commercial purposes.

### Subject Area and Field{.unnumbered}

```{r subject_data}
# Load hybrid article-level dataset and journal classification
# see
jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))
# article level data set, see data-raw/hybrid_jns_articles.R
hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
subject_hybrid <- left_join(hybrid_df, jn_subjects, by = "issn") %>%
  select(doi, issn, top_level, subject_area)

# Number of Journals (full and fractional counting)
subject_area_all_jn <- subject_hybrid %>% 
  group_by(issn, subject_area) %>%
  summarise(n = n()) %>% 
  mutate(fractions = n / sum(n)) %>% 
  ungroup() %>% 
  group_by(subject_area) %>% 
  summarise(journals_n = n_distinct(issn),
            journals_frac = sum(fractions)) %>% 
  arrange(desc(journals_frac))
# Number of articles (full and fractional counting)
subject_area_all_article <- subject_hybrid %>% 
  group_by(doi, subject_area) %>%
  summarise(n = n()) %>% 
  mutate(fractions = n / sum(n)) %>% 
  ungroup() %>% 
  group_by(subject_area) %>% 
  summarise(articles_n = n_distinct(doi),
            articles_frac = sum(fractions)) %>% 
  arrange(desc(articles_frac))
```

Next, we present disciplinary variations in hybrid OA. Table \ref{tab:subject_area_table} presents the high-level findings by AJSC subject area. In the five-year period we studied, the physical sciences (`r filter(subject_area_all_jn, subject_area == "Physical Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Physical Sciences") %>% pull(articles_n)` OA articles) and health sciences (`r filter(subject_area_all_jn, subject_area == "Health Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Health Sciences") %>% pull(articles_n)` OA articles) were the most prevalent in terms of hybrid journals with at least one OA article.  In terms of article volume, most OA articles were published in hybrid journals belonging to the life sciences (`r filter(subject_area_all_jn, subject_area == "Life Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Life Sciences") %>% pull(articles_n)` OA articles), while social sciences  (`r filter(subject_area_all_jn, subject_area == "Social Sciences") %>% pull(journals_n)`  hybrid journals, `r filter(subject_area_all_article, subject_area == "Social Sciences") %>% pull(articles_n)` articles) journals published the lowest number of OA articles. However, it is important to note that there is a large overlap between the journals’ subject areas. Therefore, Table  \ref{tab:subject_area_table} also assigns hybrid journals and OA articles fractionally to a subject area and compares it to the full counting, where journals assigned more than one subject in the AJSC classification were counted once for each field. For the remaining results, we only present the full counting.

```{r subject_area_table}
# join
inner_join(subject_area_all_jn, subject_area_all_article, by = "subject_area") %>%
  arrange(subject_area) %>%
  select(1, journals_n, articles_n, journals_frac, articles_frac) %>%
  filter(subject_area != "Multidisciplinary", !is.na(subject_area)) %>% 
  mutate(subject_area = stringr::str_to_sentence(subject_area)) %>%
  kbl(booktabs = TRUE, col.names = c("Subject area", "Journals", "OA articles", "Journals", "OA articles"),
      caption = "Subject area overview using full and fractional counting. Table reports the number of Elsevier hybrid journals with at least one open access (OA) article, and OA articles published from 2015-2019. One multidisciplinary with 61 OA articles and 11 journals with 37 OA articles that  could not be matched to the Scopus Source Title list used to obtain ASJC codes are not included.",
      format.args = list(big.mark = ",")) %>%
  add_header_above(c(" " = 1, "Full counting" = 2, "Fractional counting" = 2)) %>%
  kable_styling() 
```

```{r subject_ind}
# see 008_subject_exploration
hybrid_subject <- readr::read_csv(here::here("data", "jn_scopus_ind_subjects.csv"))
```

```{r}
subject_all_years <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(oa),
            n = sum(articles)) %>%
  ungroup() %>%
  group_by(description) %>%
  filter(subject_area != "Multidisciplinary")

median_all <- subject_all_years %>% 
  ungroup() %>% 
  distinct(issn, .keep_all = TRUE) %>% 
  mutate(prop = n_oa /n) %>% 
  .$prop %>% 
  median()
# rank
oa_per_subject_rank <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(oa),
            n = sum(articles)) %>%
  mutate(prop = n_oa / n) %>%
  ungroup() %>% 
  group_by(description) %>%
  summarise(median_prop = median(prop)) %>%
  arrange(desc(median_prop))
```

Figure \ref{fig:oa_sub_uptake} presents the journals' five-year OA uptake grouped by subject area and field and in comparison to the overall median (Mdn= `r median_all * 100`; excluding one journal coded as multidisciplinary). The box plot reveals some variation between the subject areas. Notably, with the exception of environmental science and earth and planetary sciences, journals from the physical sciences show a lower median proportion of hybrid OA articles, whereas most life sciences and social sciences journals ranked above average. Figure X also shows large variations among the 26 subject fields, with the highest OA rates in immunology and microbiology (Mdn=`r filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop)*100`), environmental science (Mdn=`r filter(oa_per_subject_rank, description == "Environmental Science") %>% pull(median_prop)*100`), and neuroscience (Mdn=`r filter(oa_per_subject_rank, description == "Neuroscience") %>% pull(median_prop)*100`), whilst  chemistry (Mdn=`r filter(oa_per_subject_rank, description == "Chemistry") %>% pull(median_prop)*100`), materials science (Mdn=`r filter(oa_per_subject_rank, description == "Materials Science") %>% pull(median_prop)*100`), and dentistry (Mdn=`r filter(oa_per_subject_rank, description == "Dentistry") %>% pull(median_prop)*100`) recorded the lowest OA rate.

```{r oa_sub_uptake, fig.cap='Open access (OA) uptake in Elsevier hybrid journals in percent grouped by subject area and field between 2015-19. Visualized as a box-and-whiskers plot limited to an OA share of 20\\%; the red line represents the overall median (2.5), jittered dots represent individual journals. Subject fields are ordered based on the median OA uptake.', out.width="90%", fig.width=9}
p_subject <- ggplot(subject_all_years, aes(fct_reorder(description, n_oa / n, .fun = median), n_oa / n)) + 
  geom_boxplot(aes(color = subject_area)) +
  geom_jitter(alpha = .05, aes(color = subject_area), size = 1) +
  facet_wrap(~subject_area, scales = "free_y") +
  geom_hline(yintercept = median_all,  colour = "red", linetype ="dashed", size = 1) +
  scale_colour_manual(NULL, values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701"),
                      guide=FALSE) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 5L),
    expand = expansion(mult = c(0, 0.05))
  ) +
  coord_flip(ylim=c(0,0.2)) +
  labs(x = NULL, y = "Five-Year Hybrid OA Uptake Rate per Journal") +
  theme_minimal_vgrid() 
p_subject
```

### Citation Impact{.unnumbered}

```{r}
hybrid_subject_19 <- hybrid_subject %>% 
  mutate(prop = oa /articles) %>%
  filter(issued_year == "2019", !is.na(percentile)
         , prop != 0) %>%
  distinct(issn, prop, percentile, snip, sjr)

cor_test_res <- cor.test(hybrid_subject_19$prop, hybrid_subject_19$percentile, method = "spearman") 
```

Furthermore, we were interested in the relationship between the journals' field-specific impact and their OA uptake. Spearman's rho correlation coefficient was used to assess the relationship between the 2019 Source Normalized Impact per Paper (SNIP) value calculated by Scopus and the journals' OA uptake that year. Considering only journals with at least one OA article, we found a weak but positive correlation between journal impact and OA uptake ($r_s$=`r as.character(round(cor_test_res$estimate,4))`, $p<0.001$), suggesting that the relationship is not very strong, based on our data (see Figure \ref{fig:sniptest}).

```{r sniptest, fig.cap="Source Normalized Impact per Paper (SNIP) versus open access (OA) uptake 2019. Each dot represents a hybrid journal with at least one published OA article in 2019. There is only a weak, positive correlation between the two variables ($r_s$=0.1679, $p<0.001$)"}
ggplot(hybrid_subject_19, aes(prop, snip)) + geom_point(alpha = 0.2) + 
  geom_smooth(method = "loess", se = TRUE) +
  coord_cartesian(xlim = c(0., 0.7), ylim = c(0.5, 6)) +
  labs(x = "Hybrid OA Uptake", y = "SNIP") +
  theme_cowplot()
```

## Invoicing for Hybrid Open Access{.unnumbered}

### Invoice Channels{.unnumbered}

```{r}
# load eligible oa articles
hybrid_volume <- readr::read_csv(here::here("data", "hybrid_articles.csv"))

# aggregate data
sponsored_articles <- hybrid_volume %>%
  mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND")) %>%
  count(issued_year, fct_source, cc_norm)
all_articles <- hybrid_volume %>%
  group_by(issued_year) %>%
  summarise(n = n())
sponsored_art <- sponsored_articles %>% 
  group_by(fct_source) %>% 
  summarise(nn = sum(n)) %>% 
  #mutate(cumsum = cumsum(nn)) %>% 
  mutate(prop = nn / sum(nn) * 100) %>%
 # mutate(cumprop = cumsum(prop)) %>%
  adorn_totals("row")
```

As can be seen from Table \ref{tab:invoicing_overview}, hybrid OA APC were most often invoiced to authors (n=`r filter(sponsored_art, fct_source == "Author") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Author") %>% pull(prop)`%) and to a lesser extent as part of agreements (n=`r filter(sponsored_art, fct_source == "Agreement") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Agreement") %>% pull(prop)`%). Interestingly, we also found a small number of cases where hybrid APCs were waived (n=`r filter(sponsored_art, fct_source == "Fee waived") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Fee waived") %>% pull(prop)`%).

```{r invoicing_overview}
sponsored_art %>% 
  kbl(booktabs = TRUE, col.names = c("Invoicing channel", "Hybrid OA articles (n)", "Percentage"),
      caption = "Invoice channels for hybrid OA in Elsevier journals between 2015-19.", digits = 1,
      format.args = list(big.mark = ",")) %>%
  kable_styling() %>%
  row_spec(4, hline_after = TRUE)
```

```{r els_waived}
els_waived_df <- hybrid_volume %>% 
  filter(oa_sponsor_type == "ElsevierWaived") 
```

Figure \ref{fig:invoiceoverview} illustrates that over the years, Elsevier has increasingly invoiced hybrid OA APCs directly to authors compared to research funders or academic consortia ("Agreement"). The share of fee-waived articles remained relatively stable, but we found different types of waivers. Around `r filter(els_waived_df, !is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)* 100`% of fee waivers were linked to an OA sponsor, which indicates that a third party paid the APC. For instance, `r filter(els_waived_df, oa_sponsor_name == "Académie des Sciences") %>% distinct(doi) %>% nrow()` waivers were linked to the French Académie des Sciences, presumably covering OA publication in its society journals for affiliated authors. The remaining `r filter(els_waived_df, is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)*100`% of waived articles did not disclose any OA sponsor. Moreover, Figure \ref{fig:invoiceoverview} compares the invoicing channels based on open content license variants. When Elsevier invoiced authors directly, most OA articles in hybrid journals were published under a non-commercial license (n=`r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC BY-NC-ND") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC BY-NC-ND") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Author") %>% pull(n) %>% sum()* 100`%), whereas most articles billed as part of agreements were licensed under the more permissive CC BY license (n=`r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC BY") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC BY") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Agreement") %>% pull(n) %>% sum() * 100`%).

```{r invoiceoverview, fig.cap="Development of fee-based open access (OA) publishing in Elsevier hybrid journals by invoice channel Colored bars represent the share of OA articles invoiced to authors, agreements, or waived, grouped by Creative Commons license variant. Grey bars show the total number of hybrid OA articles published in Elsevier journals from 2015-2019."}
# plot
sponsored_articles %>%
  mutate(fct_source = stringr::str_to_title(fct_source)) %>%
  mutate(fct_source = ordered(fct_source, levels = c("Author", "Agreement", "Fee Waived", "Other"))) %>%
  ggplot(aes(x = gsub("20", "'", issued_year), y = n), alpha = 0.8) +
  geom_bar(
    data = all_articles,
    color = "transparent",
    stat = "identity",
    fill = "#b3b3b3a0"
  ) +
  geom_bar(aes(fill = cc_norm), color = "white", stat = "identity", alpha = 0.8, size = 0.25,
           position = position_stack(reverse = T)) +
  facet_wrap( ~ fct_source, nrow = 1) +
  scale_fill_manual(values = 
                      c(`CC BY` = "#B52141",
                        `CC BY-NC-ND` = "#0093c7"),
                    name = "CC License Variant") +
  labs(x = "Publication Year", y = "Hybrid OA Articles") + 
  scale_y_continuous( 
                     labels = scales::number_format(big.mark = ","),
                      expand = expansion(mult = c(0, 0.05))) +
  theme_minimal_hgrid() +
  theme(legend.position="top", legend.justification = "right")
```

We also observed large differences in OA invoicing among subject fields. Table \ref{tab:invocing_by_subject} shows the number of OA articles by subject field and invoice channel. For articles in nursing, decision dciences, and pharmacology, toxicology and pharmaceutics, Elsevier predominantly invoiced authors, whereas most energy and chemical engineering articles were invoiced through agreements. Likewise, the majority of articles in materials science, chemistry and physics and astronomy were not invoiced to authors but facilitated through agreements or waived. The large share of waived APCs in physics and astronomy can be attributed to a single 2015 issue of Nuclear and Particle Physics Proceedings, which offered sponsored OA.

```{r invocing_by_subject, results='asis'}
hybrid_jn_volume <- hybrid_oa_volume %>%
  group_by(issn) %>% 
  summarise(n_oa = sum(oa))

jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))

hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
funder_shares <- hybrid_df %>%
   mutate(oa_sponsor_type = ifelse(!oa_sponsor_type %in% c("FundingBody", "Author", "ElsevierWaived"), "Other", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  group_by(oa_sponsor_type, issn) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  left_join(hybrid_jn_volume, by = "issn") %>%
  mutate(prop = n_oa_type / n_oa) %>%
  left_join(jn_subjects, by = "issn") %>%
  ungroup() %>%
  group_by(top_level, oa_sponsor_type) %>%
  summarise(oa_type = sum(n_oa_type)) %>%
  mutate(prop = oa_type / sum(oa_type))
# overall props
oa_sponsor_type_all <- hybrid_df %>% 
  #exclude multidisciplinary
  filter(issn != "2095-9273") %>%
  count(oa_sponsor_type) %>% 
  mutate(prop = n / sum(n))
subject_all <- funder_shares %>%
  group_by(top_level) %>%
  summarise(oa = sum(oa_type))

funder_overview_subject <- left_join(funder_shares, subject_all, by = "top_level") %>%
  tidyr::pivot_wider(id_cols = c("top_level", "oa"), names_from = oa_sponsor_type, values_from = prop) %>%
  ungroup() %>% 
  # remove na and multidisciplinary 
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  relocate(Agreement, .after = Author) %>%
  arrange(desc(Author))

make_pal <- scales::col_numeric(
  palette  = c("white", "#fcde9c","#faa476","#f0746e","#e34f6f","#dc3977"),
  na.color = "#808080",
  domain = NULL,
  alpha = FALSE,
  reverse = FALSE
)

funder_overview_subject %>%
  mutate_at(3:6, function(x){
    round(x*100,0)
  }) %>%
  mutate_at(3:6, function(x){
    cell_spec(x, background = make_pal(x))}) %>%
  kbl(booktabs = TRUE, "latex", format.args = list(big.mark = ","), digits = 1, escape = FALSE,
                  caption = "Invoice channels by discipline 2015-2019.",
                  position = "H",
      col.names = c("Subject", "OA articles", "Author", "Agreement", "Fee waived", "Other"),
       linesep = "",
      align = "lrrrrr") %>%
  add_header_above(c(" " = 2, "Invoice channel (in %)" = 4)) %>%
  kable_styling(latex_options = c("scale_down"))
```

### Invoice Recipients{.unnumbered}

```{r}
hybrid_invoice <- hybrid_volume %>%
  filter(oa_sponsor_type == "FundingBody") %>%
  left_join(curated_invoicing_df, by = c("oa_sponsor_name")) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND"))
```

```{r license_count}
sponsor_df <- hybrid_invoice %>% 
#  filter(!is.na(Country)) %>% 
  # manual changes
  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "European Union", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf))
```

Elsevier’s data offer more insight into invoicing. Overall, we identified 62 sponsoring institutions and organizations that received invoices for hybrid publishing fees as part of an agreement with Elsevier. We found that, by a large margin, most invoices were issued to UK-based research funders and institutions (n=`r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%), followed by the Netherlands (n=`r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%) and the European Union (n=`r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%; EU). Figure \ref{fig:invoice_sponsor_country} provides a closer look at the geographical distribution by license prevalence, highlighting the dominance of CC -BY licensed articles invoiced to sponsoring institutions in the UK, the United States, and, to a much lesser extent, the Netherlands. In contrast, the majority of articles invoiced to sponsoring institutions from Norway or representing the EU were published under a non-commercial license.

```{r invoice_sponsor_country, fig.cap="Number of centrally invoiced OA articles in Elsevier hybrid journals by country of recipient. Colored bars represent the Creative Commons (CC) license variant under which the article was published between 2015 and 2019."}
sponsor_df %>%
  group_by(country_factor, cc_norm) %>% 
  summarise(n = n_distinct(doi)) %>%
  ggplot(aes(fct_rev(country_factor), n, fill = cc_norm)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE),
            color = "white", alpha = 0.8, size = 0.25) +
  coord_flip() +
  scale_fill_manual(values = 
                      c(`CC BY` = "#B52141",
                        `CC BY-NC-ND` = "#0093c7"),
                    name = "CC License Variant") +
  labs(x = "OA Sponsor by Country", y = "Hybrid OA Articles (Agreement)") +
  scale_y_continuous(labels = scales::number_format(big.mark = ","),
                      expand = expansion(mult = c(0, 0.05)),
                     breaks = seq(0,15000, 2000)) +
  theme_minimal_vgrid() +
  theme(legend.position="top", legend.justification = "right")
```

Figure \ref{fig:sponsor_sector} shows the yearly distribution of central invoiced articles by year and institutional sector. While UK-based open access invoices were mainly addressed to discipline-specific governmental and non-profit research funders,  the situation is different in the Netherlands and Norway where national consortia of institutions belonging to the sector of higher education (e.g. universities) had transformative agreements with Elsevier. In 2019, Elsevier also launched similar agreements in other countries with lower publication output including Hungary and Poland. Besides, we found that invoiced funding bodies from the UK and the US mainly represent discipline-specific funders, while the remaining funding bodies from other countries are focused on a broad variety of disciplines (see Table \ref{tab:subject_profile}).

```{r sponsor_sector, fig.cap=' Development of centrally invoiced OA articles in hybrid journals by country. Colored bars represent the institutional sector OA funding bodies belong to.'}
sponsor_df %>%
  filter(!is.na(Sector)) %>%
  count(issued_year, Sector, country_factor) %>%
  mutate(Sector = stringr::str_to_title(Sector)) %>% 
  mutate(sector_level = factor(Sector)) %>%
  mutate(sector_level = fct_relevel(sector_level, "Government", "Higher Education", "Private Non-Profit", "Other")) %>%
  ggplot(aes(gsub("20", "'", issued_year), n, fill = sector_level)) + 
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE),
            color = "white", alpha = 0.8, size = 0.25) +
  facet_wrap(~country_factor) +
  scale_fill_manual("Sector", values = c("#65BADA", "#00A757", "#E5BA52", "grey50")) +
  scale_y_continuous(labels = scales::number_format(big.mark = " "),
                      expand = expansion(mult = c(0, 0.05))) +
  theme_minimal_hgrid() +
  theme(legend.position="top", legend.justification = "right",
        panel.spacing = unit(2, "lines")) +
  guides(fill = guide_legend(nrow = 2)) +
  labs(x = "Publication Year", y = "Hybrid OA Articles (Agreement)")
```

```{r}
sponsor_country <- hybrid_invoice %>%
  filter(!is.na(`Primary research focus`)) %>%
  # manual changes
  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "European Union", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf)) %>%
  mutate(`Primary research focus` = ifelse(`Primary research focus` == "SSH", "SSH", stringr::str_to_sentence(`Primary research focus`))) %>%
  group_by(country_factor, `Primary research focus`) %>%
  summarise(n = n_distinct(doi)) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n), 2) * 100)
```

```{r subject_profile}
sponsor_country %>%
  ungroup() %>%
  pivot_wider(id_cols = c("country_factor"), names_from = `Primary research focus`, values_from = prop) %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  janitor::adorn_totals(c("col", "row")) %>%
  mutate_if(is.numeric, ~paste0(., "%")) %>%
  rename(`Country/Subject area` = country_factor) %>%
  kableExtra::kbl(booktabs = TRUE,
                  caption = " Relative subject profile of centrally invoiced OA funding bodies.", align = c("l|rrrrr|r")) %>%
    row_spec(6:7, hline_after = TRUE) %>%
   kable_styling(latex_options = c("scale_down"))
```

### Institutional ranking{.unnumbered}

```{r open_apc_check}
hybrid_invoice <- hybrid_invoice %>%
  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
```

Table \ref{tab:sponsor_top_10} presents the top ten of `r hybrid_invoice %>% filter(!is.na(oa_sponsor_cleaned)) %>% distinct(oa_sponsor_cleaned) %>% nrow()` invoiced institutions and organizations in our sample. Together, they accounted for around 80% of centrally invoiced APCs. The table also highlights the number of distinct journals and the share of CC BY licensed articles. Notably, UK-based research funders and the US-based Bill and Melinda Gates Foundation mainly sponsored articles published under a CC BY license. In contrast, articles funded by VSNU, the European Research Council, and Norway Institutes had a much lower proportion of CC BY.

```{r, sponsor_top_10}
hybrid_invoice_tt <- hybrid_invoice %>%
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_cleaned), "Other", oa_sponsor_cleaned)) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND")) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
hybrid_invoice_agg_general <- hybrid_invoice_tt %>%
  group_by(oa_sponsor) %>%
  summarise(journals = n_distinct(issn),
            articles = n_distinct(doi)) %>%
  mutate(prop = articles / sum(articles))
hybrid_invoice_agg_cc <- hybrid_invoice_tt %>%
  ungroup() %>%
  filter(cc_norm == "CC BY") %>%
  group_by(oa_sponsor) %>%
  summarise(cc_by = n_distinct(doi))
hybrid_invoice_agg_oapc <- hybrid_invoice_tt %>%
  filter(in_open_apc == TRUE) %>%
  group_by(oa_sponsor) %>%
  summarise(in_open_apc =  n_distinct(doi))
ind_sponsor_df <-
  left_join(hybrid_invoice_agg_general, hybrid_invoice_agg_cc, by = "oa_sponsor") %>%
  left_join(hybrid_invoice_agg_oapc, by = "oa_sponsor") %>%
  mutate(
    cc_by = cc_by / articles,
    in_open_apc = ifelse(is.na(in_open_apc), 0, in_open_apc),
    in_open_apc = in_open_apc / articles
  ) %>%
  mutate(oa_sponsor = fct_reorder(oa_sponsor, articles)) %>%
  mutate(oa_sponsor = forcats::fct_relevel(oa_sponsor, "Other", after = 0L)) %>% arrange(desc(oa_sponsor))
# overall
overall_ind_sponsor_df <-
  tibble(
    oa_sponsor = "All",
    journals = length(unique(hybrid_invoice$issn)),
    articles = length(unique(hybrid_invoice$doi)),
    prop = 1,
    cc_by = filter(hybrid_invoice, cc_norm == "CC BY") %>% nrow() / length(unique(hybrid_invoice$doi)),
    in_open_apc = filter(hybrid_invoice, in_open_apc == TRUE) %>% nrow() / length(unique(hybrid_invoice$doi))
  )

ind_sponsor_df %>%
  bind_rows(overall_ind_sponsor_df) %>%
  mutate(oa_sponsor = gsub("&", "and", oa_sponsor)) %>%
  mutate_at(4:6, function(x){
    round(x*100,0)
  }) %>%
  mutate_at(5:6, function(x){
    cell_spec(x, background = make_pal(x))}) %>%
  kbl("latex", booktabs = TRUE, escape = FALSE,
      col.names = c("Hybrid OA sponsors", "Journals", "Articles", "\\%", "CC BY", "OAPC"),
      align = c("lrrrrr"),
      linesep = "",
      caption = "Top 10 Elsevier hybrid OA sponsors by centrally invoiced articles 2015-2019.  The table furthermore presents the number of Elsevier hybrid journals with at least one invoiced article and two indicators of compliance: proportion of articles provided under a CC BY license and where spending was disclosed through the Open APC Initiative (OAPC). VSNU and Norway Institutes paid for OA through transformative agreements according to the ESAC Transformative Agreement Registry.") %>%
  row_spec(11:12, hline_after = TRUE) %>%
  add_header_above(c(" " = 4, "Compliance (in %)" = 2)) %>%
   kable_styling(latex_options = c("scale_down"))
```

The table also highlights the proportion of APC funding publicly disclosed through the Open APC initiative, suggesting that funders with a large CC BY share were also more willing to share their spending publicly. On the other hand, APC funding activities through transformative agreements were not publicly shared.

Figure \ref{fig:field_funder_matrix} shows the subject profile of OA funders relative to the journals’ subject fields. As expected, the figure shows that subject-specific funders tended to focus on a limited set of subject fields. In contrast, national consortia, such as VSNU and Norwegian Institutes (NO), financed OA articles  for a broader range of fields. However, within these consortia, the subject distribution also varied. VSNU OA spending, for example, had a strong focus on articles relating to Medicine, Environmental Science, and Energy, presumably reflecting disciplines with generally strong publication output at Dutch universities. Norwegian universities mostly published hybrid OA in Engineering, Environmental Science and Agricultural and Biological Sciences journals.

```{r}
#short funder labels
funder_codes <- tribble(
  ~oa_sponsor, ~oa_sponsor_short,
  "Engineering and Physical Sciences Research Council", "EPSRC",
  "VSNU", "VSNU",
  "Wellcome Trust", "Wellcome",
  "European Research Council", "ERC",
  "Medical Research Council", "MRC",
  "Natural Environment Research Council", "NERC",
  "Biotechnology and Biological Sciences Research Council", "BBSRC",
  "Bill & Melinda Gates Foundation", "Gates",
  "Economic and Social Research Council", "ESRC",
  "Norway Institutes", "NO",
  "Other", "Other")
```

```{r field_funder_matrix, out.width="90%", fig.width=8, fig.height=12}
field_funder <- hybrid_df %>%
  filter(oa_sponsor_type == "FundingBody") %>%
  left_join(funder_codes, by = c("oa_sponsor_name" = "oa_sponsor")) %>%
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_short), "Other", oa_sponsor_short)) %>%
  mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  left_join(jn_subjects, by = "issn") %>%
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  group_by(oa_sponsor, top_level) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
  mutate(oa_sponsor = fct_relevel(oa_sponsor, "Other", after = Inf))
# plot
pp <- ggplot(field_funder,
       aes(
         x = factor(oa_sponsor),
         y = reorder(top_level, desc(top_level)),
         colour = prop,
         size = n_oa_type
       )) +
  geom_point() +
  scale_color_viridis_c(
    labels = scales::percent_format(accuracy = 5L),
    option = "magma",
    name = "Percentage per Funder",
    breaks =  scales::extended_breaks(),
    begin = 0.15,
    end = 0.85
  ) +
  scale_size("OA Articles") +
  scale_x_discrete(position = "top") +
  labs(x = NULL, y = NULL) +
  theme_minimal_grid(font_size = 12) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()
  ) +
  theme(legend.position = "top", legend.justification = "right") +
  theme(legend.direction = "horizontal", 
        legend.box = "horizontal")
ggsave(here::here("figure", "funder_subject.png"), pp, dpi = 300, width = 9.5, height = 9)
```

\begin{figure}[H]

{\centering \includegraphics[width=0.95\textwidth]{../figure/funder_subject.png}

}

\caption{Top 10 Elsevier hybrid OA institutional invoice recipients in terms of articles published by journal subject. EPSRC, Engineering and Physical Sciences Research Council; VSNU, VSNU; Wellcome, Wellcome Trust; ERC, European Research Council; ERC, European Research Council; MRC, Medical Research Council; NERC, Natural Environment Research Council; BBSRC, Biotechnology and Biological Sciences Research Council; Gates, Bill and Melinda Gates Foundation; ESRC, Economic and Social Research Council; NO, Norway Institutes. }\label{fig:field_funder_matrix}
\end{figure}