### 2021-01-31 analysis/011_scopus_subject.Rmd
scopus whole counting

---
title: "011_scpopus_subject_all.Rmd"
output: 
  github_document: 
    df_print: kable
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r}
library(tidyverse)
library(janitor)
```

read in scopus journal indicators

```{r}
scopus_19 <- readxl::read_excel(here::here("data", "scopus_jn.xlsx"), sheet = 2,
                                .name_repair = make_clean_names)
ajsc_mapped <- read_csv(here::here("data", "asjc_mapped.csv")) %>%
  mutate(top_level_code = as.character(top_level_code))
```

normalize scopus jn indicators

```{r}
scopus_norm <- scopus_19 %>%
  mutate(top_level_code = substr(as.character(scopus_asjc_code_sub_subject_area), 1, 2)) %>%
  inner_join(ajsc_mapped, by = "top_level_code")
```

## All

```{r}  
scopus_norm %>%
  distinct(subject_area, scopus_source_id, scholarly_output) %>%
  group_by(subject_area) %>% 
  summarise(n = sum(scholarly_output)) %>%
  mutate(prop = n / sum(n)) %>%
  arrange(desc(n))
```

## Elsevier

```{r}  
scopus_norm %>%
  filter(publisher == "Elsevier") %>%
  distinct(subject_area, scopus_source_id, scholarly_output) %>%
  group_by(subject_area) %>% 
  summarise(n = sum(scholarly_output)) %>%
  mutate(prop = n / sum(n))
```

By oa

```{r}  
scopus_norm %>%
  filter(publisher == "Elsevier") %>%
  distinct(subject_area, scopus_source_id, scholarly_output, open_access) %>%
  group_by(subject_area, open_access) %>% 
  summarise(n = sum(scholarly_output))
```