### 2021-05-19 analysis/manuscript.Rmd
sv tmp

---
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: apa6.sty
csl: apa.csl
bibliography: references.bib
---

```{r knitr_options, , echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300,
  fig.pos = "H", 
  out.extra = "",
  dev= c("png", "cairo_ps")
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(formatC(x, big.mark = ",", format = "f", digits = 1, drop0trailing = TRUE))
    } else{
      return(x)
    }
  }
)
```

```{r libraries}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(gt)
library(janitor)
library(kableExtra)
library(stringr)
```
# Introduction

The rise of open access (OA) has added complexity to
scholarly publishing, particularly concerning transparency of economic
dimensions. Financial transparency in journal publishing has long been
lacking because Big Deal subscription contracts between academic
institutions and large academic publishers usually include
confidentiality clauses that prohibit publicly sharing
the scope or pricing of such agreements [@Frazier_2001; @Bergstrom_2014; @Larivi_re_2015]. Further, the distributed payment processes
to access and publish scholarly articles add to the opacity of
the financial flows of scholarly publishing [@Lawson_2016] as they can
involve several actors simultaneously, including authors, research
institutions, libraries, and funders.

The problem with this lack of transparency becomes more apparent in the
case of hybrid OA. In this OA business model, individual articles can be
made openly available by paying an article processing charge
(APC), while the journal as a whole remains subscription-access. Hence,
hybrid OA was originally introduced as a low-risk transitional model
that allows journals to gradually convert to full OA and reduce
subscription costs as the OA uptake increases [@Prosser_2003]. Many
large subscription publishers have since incorporated hybrid OA, but there have been concerns that the hybrid model allows for double dipping---receiving two payments for one article, the APC and
subscription fees [@Shieber_2009; @Prosser_2015]. While publishers
assure they adjust their pricing
[@Mittermaier_2015], without transparency around the uptake of hybrid OA
and both revenue streams, such claims are impossible to evaluate. This is especially concerning because hybrid OA has recently gained popularity after a decade of only slow uptake [@Bj_rk_2012]. The main drivers of this development have been research institutions and funders that implemented  OA policies, allocated OA funds, and established
formal agreements with publishers that allow
affiliated authors to publish OA free-of-charge [@Laakso_2016].

The recent introduction of transformative agreements, an evolving
concept describing contracts that shift library spending from subscriptions to OA
[@Borrego_2020; @Hinchliffe_2019], might perpetuate the lack of transparency because pricing is often based on traditional subscription costs. Hence, the demand for
publisher-provided data has increased. More
transparency about hybrid OA uptake and funding could  facilitate the assessment and adjustment of publisher contracts [@Schimmer_2015], enhance OA mandate compliance monitoring, and avoid double-dipping [@Larivi_re_2018]. However, previous studies have noted an absence of transparency in hybrid OA publishing and a considerable lack of
publicly available and standardized data [@Laakso_2016; @Lawson_2015; @Pinfield_2016].

This paper investigates to what extent publisher-provided
data can increase the transparency around hybrid OA by taking Elsevier,
the largest scholarly journal publisher [@Larivi_re_2015], as the starting point for an empirical
analysis of hybrid OA uptake and invoicing. Our approach is based on
openly available data sources and combines scholarly metadata
from Crossref, a DOI registration agency, with
 APC invoicing data from Elsevier. The
compiled data enable us to analyze the 
hybrid OA uptake in Elsevier journals between  2015 and 2019 by license and subject, and to compare it with
Elsevier's full and delayed OA program. Further, we determine whether
hybrid APCs were waived or charged to authors or as part of publishing agreements. In the latter case, we 
examine the invoiced academic consortia and research funders. As such, our findings have implications for research and OA
policy and highlight the potential of publisher-provided data for
large-scale, instantaneous analysis of hybrid OA uptake and invoicing.

# Background

## Prevalence / uptake studies

Since there is no standardized way to measure the uptake of hybrid OA, estimates can vary greatly depending on how uptake is operationalized--for example, through the number of hybrid journals with at least one OA article, the average number of OA articles per hybrid journal, or the share of OA articles compared to subscription articles.

In the first article-level study on hybrid OA, @Laakso_2016
reported that between 2007 and 2013, the five largest
publishers--Elsevier, Sage, Springer, Taylor \& Francis, and
Wiley--recorded growing numbers of hybrid journals that coincided with a
twenty-fold increase in hybrid OA (from 666 to 13,994). As a
result, hybrid journals with at least one OA article more than doubled
in number (1,082 in 2009, 2,714 in 2013) but fell by 31\% relative to
the total number of hybrid journals. Laakso and Björk's
(2016) exploratory approach yielded a comprehensive overview tracking the growth of hybrid
OA over time and across multiple publishers, but due to the amount of
manual data collection and cleaning it
is unsuitable for repeated use. Minimizing such manual tasks,
Nelson \& Eggett (2017) narrowed their focus to one publisher--the
American Chemical Society (ACS)--and requested a list of all hybrid OA
articles published from 2006-2011 from the ACS. To assess the
hybrid OA uptake, the authors aggregated and compared the number of OA
(n=814) and subscription articles (n=27,621), finding that over the five
years, 2.9\% of ACS articles were published as hybrid OA.

Contrary to the previous two studies, @Kirkman_2018 took a research
funder as a starting point, reporting a 26\% share of hybrid OA for
articles funded by the Australian National Health and Medical Research
Council from 2012-2014 (816 of 3,190). Kirkman collected research
articles using the Funding Agency and Funding Text search fields on Web
of Science, and classified articles as OA when they were freely
available on the publisher website, and further distinguished hybrid OA
articles based on journal-level information from the publisher website,
such as hybrid OA journal and APC lists. Another approach is presented
by @P_l_nen_2020, who used current research information system
(CRIS) data---institutional publication data---to study the extent of OA
among Finnish research publications from 2016-2017 (n=34,507 research
articles). As part of the analysis, the authors identified hybrid OA
articles through a dedicated OA status metadata field that has been
mandatory to report in Finland since 2016 and found a 7\% share of
hybrid OA. However, the drawbacks of Kirkman's (2018) and Pölönen et
al.'s (2020) approaches are that such data sources might not be openly
available or, regarding CRIS data, might not be available at all because
not all countries maintain comprehensive institutional publication data.

Other studies assessed the share of (hybrid) OA for the entire corpus of research articles. Using DOI-assigned research articles indexed in the Web of Science Core Collection
 as a benchmark, @Mart_n_Mart_n_2018 investigated how
many research articles from 2009 and 2014 (n=2,610,305 in total) are freely available through
Google Scholar. After identifying OA articles through licensing
information from Crossref, the authors estimated the share of hybrid OA
at around 0.5\% in 2009 and 1.5\% in 2014. Similarly, in a study assessing the OA level among all scholarly articles and that experienced by Unpaywall users, Piwowar et
al. (2018) obtained license information through Crossref and web-scraping. The study found 3.6\% of 100,000 random Crossref DOIs were hybrid OA
and pointed towards a growing trend in recent years that recorded a
hybrid OA share of 9.4\% of all articles published in 2015.

Since its launch, Unpaywall has become widely used in bibliometric research and rankings [@Robinson_Garcia_2020; @Huang_2020]. However, the lack of standardized and comprehensive publisher-provided data has required several updates to Unpaywall to improve hybrid OA identification and differentiation from  delayed OA [@Piwowar_2019; @Unpaywall_oa_license; @Unpaywall_types], illustrating ongoing challenges in tracking and comparing hybrid OA prevalence over time. After changes in Unpaywall's approach to detect hybrid OA, Piwowar et al. (2019) estimated a hybrid OA share of 4\% amongst all articles published by July 2019.

## Financial studies

Pinfield et al.'s (2016) analysis of APC payment records provided by 23
United Kingdom (UK) higher education institutions revealed a sharp increase in central
payments from 2007-2013, which was largely attributed to the
introduction of block grants by Research Councils UK (RCUK) and
non-compliance sanctions by the Wellcome Trust. Moreover, the study
showed that OA fees were paid almost exclusively through block grants (92\%), and only a small number of APCs were paid
through internal funding (7\%). In contrast, a recent Springer Nature
survey found that authors draw on a range of funding sources to cover OA
fees, such as dedicated institutional OA funds, block grants, OA
agreements, or research grants (Monaghan et al., 2020). Most hybrid OA
authors were supported through dedicated institutional OA funds (43\%,
excluding block grants) and OA agreements with Springer Nature (34\%).
The differences between these two studies might be due to Monaghan et al.'s
(2020) focus on Springer Nature articles or due to} different policy and funding arrangements---considering the introduction of OA
agreements since  @Pinfield_2016 and Monaghan et al.'s
(2020) more regionally diverse sample.

Regional and policy differences in APC payments also came to light in a
study by @Jahn_2016 that analyzed APC records from 30 German
higher education and research institutions, the Austrian Science Fund (FWF), Jisc, and the Wellcome
Trust. In particular, the study revealed large differences in the amount
of hybrid OA funded from 2014-2015. Whilst hybrid OA accounted for less
than 1\% of APCs paid by German institutions (23 of 3,846), the three
non-German research funders recorded a hybrid share of 75\% (11,533 of
15,779). According to @Jahn_2016, this could point towards
differences in science policy, such as hybrid OA being supported by the
three non-German research funders but not by Germany's largest national
funder, the German Research Foundation (DFG). Another possibility is that German hybrid OA fees were
paid from budgets not reported to the Open APC initiative, a crowd-sourcing effort [@Pieper_2018] from where
the authors acquired data. Among these unreported funds
are research grants and research unit budgets, which author surveys
identified as APC funding sources [@Monaghan_2020; @Graaf_2017]. As such, Jahn and Tullney's (2016) findings could reflect
the complexities and potential limitations of institutional OA spending
data that @Pinfield_2016 and @Monaghan_2020 attributed
to incomplete or missing records.

In recent years, national consortia in Europe have negotiated publishing agreements covering hybrid OA fees for affiliated authors (Borrego et al., 2020). While these have improved invoicing workflows, internal assessments of hybrid OA uptake and invoicing remain challenging because transparent and comparative data have largely remained absent [@Marques_2020]. Such publicly available information might be scarce because publishers and institutions lack or withhold such data, or due to confidentiality clauses [@Marques_2020; @Monaghan_2020]. Several European consortia have collaborated through the Efficiency and Standards for Article Charges (ESAC) Initiative and Knowledge Exchange to standardize metadata requirements across publishers as great variation has been found between different countries and publishers [@Marques_2019].

## Research questions and aims

In this paper, we focus on Elsevier, a prominent example in recent hybrid OA uptake and financial studies [@Pinfield_2016; @@Laakso_2016]. Elsevier's OA portfolio presents the challenges in examining hybrid OA described above. For instance, distinguishing between different OA types. Elsevier supports delayed [@Els_Archive], hybrid, and full OA, including so-called mirror journals---full OA counterparts of hybrid journals addressing OA policies opposed to hybrid OA [@Harrison_2019]. Further, Elsevier processes APC invoices through various channels, such as agreements with research funders and library consortia or, in the absence thereof, the authors (Elsevier,
n.d.-a, n.d.-c). Surprisingly, we found Elsevier article-level metadata embedded in XML full-texts indicating the articles' OA status and invoicing.  Leveraging this publicly available data, we 
address the lack of transparency around hybrid OA noted in previous studies.
In particular, we use this novel approach to answer the following
questions:

* What was the uptake of Elsevier's hybrid OA publishing option between 2015 and 2019?
* Through which channels were hybrid APCs invoiced, and who were the recipients?

# Methodology

For this study, we collected data relating to Elsevier's hybrid OA
option by drawing on multiple freely available data sources. We
identified Elsevier hybrid journals through Elsevier's APC list and
supplemented our sample with Crossref metadata and text-mined invoicing
data to investigate the invoicing of immediate OA articles provided
under a Creative Commons (CC) license in subscription-based journals. Figure
\ref{fig:workflow} visualizes the automated workflow we used to collect
data from Elsevier and Crossref.

\begin{figure}[H]

{\centering \includegraphics[width=0.95\textwidth]{../figure/els_flow.pdf}

}

\caption{Data Collection Workflow to Obtain Article-Level OA and Invoicing Data.
}\label{fig:workflow}
\end{figure}

First, we identified hybrid journals through an Elsevier APC price list
from May 2020 provided by @Matthias_2020. We excluded journals that transitioned from hybrid to full OA and
reverse-flip journals that flipped 
from full to hybrid OA [@Matthias_2019]. Overall, we identified 1,970 unique hybrid journals that published between 2015 and 2019.

Next, we used an openly available Crossref database snapshot [@Crossref_2020], which contains all Crossref records registered until March 2020,
to calculate the journals' combined article volume for the five-year
period from 2015-2019. We only included articles published in regular
issues aside from supplements containing conference contributions like
meeting abstracts, indicated by non-numeric pagination. We excluded
non-scholarly journal content, such as the table of contents, following Unpaywall's paratext recognition approach
 [@Unpaywall_para], which we expanded to include patterns indicating
corrections. Furthermore, we categorized the articles by subject
according to the All Science Journal Classification code (ASJC) and
added citation indicators from the Scopus Source Title list (version
June 2020) based on the journal they were published in.

We identified OA articles
through CC licenses in Crossref metadata records (Hendricks et al., 2020) and then downloaded the XML
version of all CC-licensed articles published in a hybrid journal. From
the XML files, we obtained the articles' OA status and invoicing information (see Table \ref{tab:els_xml}). We determined whether articles were immediate or delayed
OA using the XML node \texttt{openArchiveArticle}, which indicates if an article was made freely available under Elsevier's Open Archive program, and measured the
uptake of hybrid OA. Moreover, we obtained the invoice channels and recipients of hybrid OA APCs. Using the \texttt{openaccessSponsorType} node, we distinguished between four invoice channels, including
invoices billed to authors, as part of publishing
agreements with funding bodies (cf. @Els_Agreements), exempted through fee waivers (e.g., in ``cases of genuine need'' or due to
society or university sponsorships, cf.~Elsevier,
n.d.-c), and other types not
specified by Elsevier. Finally, when hybrid APCs were invoiced as part
of agreements, we identified invoice recipients through the
\texttt{openaccessSponsorName} node. It is worth noting that while this article focuses on OA articles published in hybrid journals, the same data is also available for articles published in full OA journals. Elsevier does not provide the invoiced APC in either case.

\begin{table}[H]
\caption{\label{tab:els_xml}Metadata in Elsevier XML Full-Texts.}
\centering
\begin{tabular}[t]{lp{10cm}}
\toprule
XML node & Description\\
\midrule
\texttt{openaccessArticle} & Was the article open
access? \\
\texttt{openaccessSponsorType} & Invoice channel.\\
\texttt{openaccessSponsorName} & Invoice recipient.\\
\texttt{openArchiveArticle} & Was open access provided through
Elsevier's Open Archive program?\tabularnewline \\
\bottomrule
\end{tabular}
\end{table}

We manually classified invoice recipients based on their institutional
sectors, countries, and primary research areas. Following the OECD's
Frascati Manual [@Frascati, p. 91], we coded for four sectors: business
enterprise, government, higher education, and private non-profit. 
Due to the low article volume,
we combined the business enterprise sector with invoice recipients Elsevier listed as "authors" and
"third-party sponsor"
 into "Others". Moreover, we categorized invoice
recipients according to the countries representing their
scope of funding and based on the following primary research areas
health sciences, life sciences, physical sciences and mathematics,
social sciences and humanities, broad (i.e., multiple
research areas), and unknown. Further, we compared Elsevier's invoicing data
with institutional spending data from the Open APC initiative.

Throughout this mostly automated data gathering and analysis process, we
used tools from the Tidyverse [@tidyverse] for the R
programming language [@r]. To allow for efficient data
manipulation and retrieval, we imported the Crossref dump to Google
BigQuery, applying the rcrossref [@rcrossref] parsers to
extract relevant metadata fields. We used crminer
[@crminer] to obtain the XML-full texts from Elsevier.

# Results{.unnumbered}

This section first presents the results of our analysis of hybrid OA
uptake in Elsevier's journal portfolio with a view to licensing,
disciplinary differences, and citation impact. Then, we present a
descriptive analysis of Elsevier's invoicing data, highlighting
licensing and disciplinary differences for invoicing channels and invoice recipients.

```{r data}
# journal-level data
els_jns <- readr::read_csv(here::here("data", "els_historic_jns.csv"))
els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
# article-level data
hybrid_oa_volume <- readr::read_csv(here::here("data", "hybrid_oa_volume.csv"))
# invoicing analysis
curated_invoicing_df <- readr::read_csv(here::here("data", "curated_invoicing_data.csv"))
# open apc data
o_apc <- readr::read_csv(here::here("data", "o_apc.csv"))
```

## Uptake of Hybrid Open Access{.unnumbered}

### Overview{.unnumbered}

```{r}
# journals by year and oa model
els_jns_count <- els_jns %>%
  group_by(year, oa_model) %>%
  count(year) %>%
  group_by(year) %>%
  mutate(prop = n / sum(n)) %>%
  mutate(oa_model_fct = fct_rev(as_factor(oa_model)))
# overall count
 els_overall <- els_jns %>% 
  group_by(oa_model) %>% 
  summarise(n = n_distinct(issn))
```

```{r, rev_flipped}
flipped_jns <- els_jns %>%
  distinct(issn, oa_model) %>%
  group_by(issn) %>%
  filter(n() > 1)
flipped_jns_ind <- els_jns %>%
  filter(issn %in% flipped_jns$issn) %>%
  group_by(issn) %>%
  slice(which.max(year)) %>%
  ungroup() %>%
  count(oa_model)
```

```{r hybrid_oa_per_year}
# journal and article metrics **per year**
hybrid_oa_per_year <- hybrid_oa_volume %>%
  group_by(issued_year) %>%
  # distinct jns with at least one article per year
  filter(oa > 0) %>%
  summarise(
      # distinct jns with at least one article per year
    n_journals = n(),
    # article metrics
    articles_n = sum(articles),
    articles_mean = mean(articles),
    articles_sd = sd(articles),
    # oa articles metrics
    oa_n = sum(oa),
    oa_mean = mean(oa),
    oa_sd = sd(oa)
    ) %>%
  mutate(oa_uptake = oa_n / articles_n)
# oa uptake per year
oa_uptake_per_year <- hybrid_oa_volume %>% 
  filter(oa > 0) %>%
  mutate(prop = oa / articles) %>%
  group_by(issued_year) %>%
  summarise(oa_uptake_mean = mean(prop),
            oa_uptake_sd = sd(prop))
# merge indicators
ind_per_year <- inner_join(hybrid_oa_per_year, oa_uptake_per_year, by = "issued_year")

# five-years article volume
all_jn_articles <- hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    articles_p_jn = sum(articles)
    ) %>%
  ungroup() %>%
  summarise(
    n_journals = n(),
    articles_n = sum(articles_p_jn),
    articles_mean = mean(articles_p_jn),
    articles_sd = sd(articles_p_jn)
  )
# five-years OA article volume
all_oa_articles <- 
  hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    oa_p_jn = sum(oa)
    ) %>%
  ungroup() %>%
  summarise(
    oa_n = sum(oa_p_jn),
    oa_mean = mean(oa_p_jn),
    oa_sd = sd(oa_p_jn)
  )
# five-years OA uptake rate
oa_uptake <- 
  hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    articles_p_jn = sum(articles),
    oa_p_jn = sum(oa)
    ) %>%
  mutate(prop = oa_p_jn / articles_p_jn) %>%
  ungroup() %>%
  summarise(oa_uptake = sum(oa_p_jn) / sum(articles_p_jn),
            oa_uptake_mean = mean(prop),
            oa_uptake_sd = sd(prop)
  )
# bringing it all together
five_years_ind <- bind_cols(issued_year = "2015-19", all_jn_articles, all_oa_articles, oa_uptake)

ind_summary_table <- ind_per_year %>%
  mutate(issued_year = as.character(issued_year)) %>% 
  bind_rows(five_years_ind) %>%
  mutate(across(contains("uptake")) * 100) %>%
  pivot_longer(n_journals:oa_uptake_sd, names_to = "indicator", values_to = "count") %>%
  pivot_wider(names_from = issued_year, values_from = count)
```

Between 2015 and 2019, `r hybrid_oa_volume %>% distinct(issn) %>% nrow()` of `r els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` (`r  hybrid_oa_volume %>% distinct(issn) %>% nrow() / (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n))*100`%) hybrid journals published at least one OA article. Together these journals published `r five_years_ind %>% pull(oa_n)` OA articles, which represents `r five_years_ind %>% pull(oa_uptake)*100`% of their total article volume (n=`r five_years_ind %>% pull(articles_n)`). Table \ref{tab:overview} presents  findings for each year and the aggregated five-year period, illustrating moderate growth in hybrid OA over time. Each year, the number of hybrid journals
with at least one OA article increased, as did the number of immediate
OA articles in these journals, which nearly doubled from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_n)` in 2015 to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_n)` in 2019. However, since the journals' total article output also grew over time, the relative share of OA articles
in hybrid journals only increased slightly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake)*100`% to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake)*100`%.

```{r overview}
table_descriptives_ind <- tribble(
  ~ind_group, ~ind, ~` `,
    "Elsevier Hybrid Journals with ≥ 1 OA article (n)", "n_journals", "Total",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "articles_n", "Total ",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "articles_mean", "Avg.",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "articles_sd", "SD",
  "OA Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "oa_n", "Total",
  "OA Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "oa_mean", "Mean",
  "OA Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "oa_sd", "SD",
  "Share of Elsevier Hybrid Journals with ≥ 1 OA article (%)", "oa_uptake", "Total",
  "Share of Elsevier Hybrid Journals with ≥ 1 OA article (%)", "oa_uptake_mean", "Mean",
  "Share of Elsevier Hybrid Journals with ≥ 1 OA article (%)", "oa_uptake_sd", "SD"
)
inner_join(table_descriptives_ind, ind_summary_table, by = c("ind" = "indicator")) %>%
  dplyr::select(-ind, -ind_group) %>%
  kableExtra::kbl(booktabs = TRUE, format.args = list(big.mark = ","), digits = 1, escape = TRUE,
                  caption = "Elsevier Hybrid OA Uptake  2015-2019.",
                  position = "H"
) %>%
  kable_styling() %>%
  pack_rows("Elsevier Hybrid Journals with ≥ 1 OA article (n)", 1, 1) %>%
  pack_rows("Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", 2, 4) %>%
  pack_rows("OA Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", 5, 7) %>%
  pack_rows("Share of Elsevier Hybrid Journals with ≥ 1 OA article (%)", 8, 10)
```

```{r}
perc_df <- hybrid_oa_volume %>%
  mutate(prop = oa / articles) %>%
  filter(oa > 0) %>%
  group_by(issued_year) %>% 
  summarize(median = median(prop),
            p_10 = quantile(prop, 0.1),
            p_95 = quantile(prop, 0.95),
            iqr = quantile(prop, 3/4) - quantile(prop, 1/4)) %>%
  ungroup()
```

Figure \ref{fig:boxuptake} presents the hybrid OA uptake by year, showing considerable variation among Elsevier's hybrid journals. The box-and-whiskers plot visualizes the median hybrid OA uptake (horizontal line inside the box), the range of the majority of journals (the box), and individual outliers (dots). Notably, the upper quartiles and whiskers stretch farther over the years, which indicates that the range of uptake among journals with an above-average proportion of OA articles increased over the years. Nevertheless, the prevalence of hybrid OA remained relatively low; in 2019, 95% of Elsevier hybrid journals published ≤`r perc_df %>% filter(issued_year == 2019) %>% pull(p_95) * 100`% of their articles OA.

```{r boxuptake, fig.cap = "Open access uptake per Elsevier hybrid journal in percent by year, visualized as box-and-whiskers plot. The Y-axis is limited to an open access share of 20\\%."}
hybrid_oa_volume %>%
  mutate(prop = oa / articles) %>%
  filter(oa > 0) %>%
ggplot(aes(factor(issued_year), prop)) + 
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) + 
  coord_cartesian(ylim=c(0,0.2)) +
  theme_minimal_hgrid() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L),
      expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Publication Year",
       y = "Hybrid OA Uptake")
```

### Share of the Total Volume published in OA{.unnumbered}

```{r oa_volume}
# see analysis/006_license.Rmd how the data was obtained

license_df <- readr::read_csv(here::here("data", "year_per_oa_type_and_license.csv"))
# all oa
all_oa <- license_df %>% 
  group_by(type) %>% 
  summarise(n = sum(articles)) %>%
  pull(n) %>%
  sum()
# all articles
all_els <- license_df %>%
  distinct(all_articles, issued_year) %>%
  pull(all_articles) %>%
  sum()
# summary by type
summary_oa_type <- 
  license_df %>% 
  group_by(type) %>% 
  summarise(n = sum(articles)) %>%
  mutate(prop = n / sum(n))
# mirror journals
mirror_jn <- readr::read_csv(here::here("data", "x_journals_volume.csv"))
```

During the same period, Elsevier published `r all_oa` OA articles in total (i.e., in hybrid and full OA journals), representing `r all_oa / all_els * 100`% of the total article volume (n=`r all_els`). Looking at Figure \ref{fig:oa_volume_fig}, it is apparent that the OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`; `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged behind that of Elsevier's Open Archive program (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`; `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) and full OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`; `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r sum(mirror_jn$all_articles)` articles published in `r length(unique(mirror_jn$container_title))` mirror journals.

```{r oa_volume_fig, fig.cap="Open Access (OA) in Elsevier journals in terms of article volume between 2015-2019. Figure A presents the aggregated OA article volume of Elsevier journals from 2015-2019 by OA type. Figure B presents the share of OA articles relative to Elsevier's overall article output by OA type and open content license from 2015-2019."}
# see analysis/006_license.Rmd how the figure was created

knitr::include_graphics(here::here("figure", "license_portfolio.png"))
```

### License Prevalence{.unnumbered}

```{r cc_by_by_type}
cc_by_by_type <- license_df %>%
  group_by(type, license_code) %>%
  summarise(n = sum(articles)) %>%
  mutate(prop = n / sum(n)) %>%
  filter(license_code == "CC BY")
```

As far as we observed, OA articles in Elsevier hybrid journals were published under two possible CC licenses--CC BY or CC BY-NC-ND. CC BY allows others to distribute, remix, adapt, and build upon the licensed work, including for commercial purposes, as long as the original author is credited. In contrast, CC BY-NC-ND is less permissive as it prohibits commercial reuse and derivatives. In the five-year period from 2015-2019, the proportion of hybrid OA articles published under a CC BY-NC-ND license marginally increased and maintained a greater share than that of CC BY (see Figure \ref{fig:oa_volume_fig}-B). Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid journals had the highest number and proportion of openly available articles licensed under CC BY (n=`r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(n)`; `r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(prop) *100`%) compared to full OA journals  (n=`r filter(cc_by_by_type, type == "full_oa") %>% pull(n)`; `r filter(cc_by_by_type, type == "full_oa") %>% pull(prop) *100`%) and Elsevier's delayed OA program (n=`r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(n)`; `r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(prop) *100`%) . Most  delayed OA articles were provided under an Elsevier user license (Els-User), which prohibits reuse for commercial purposes.

### Subject Area and Field{.unnumbered}

```{r subject_data}
# Load hybrid article-level dataset and journal classification
# see
jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))
# article level data set, see data-raw/hybrid_jns_articles.R
hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
subject_hybrid <- left_join(hybrid_df, jn_subjects, by = "issn") %>%
  select(doi, issn, top_level, subject_area)

# Number of Journals (full and fractional counting)
subject_area_all_jn <- subject_hybrid %>% 
  group_by(issn, subject_area) %>%
  summarise(n = n()) %>% 
  mutate(fractions = n / sum(n)) %>% 
  ungroup() %>% 
  group_by(subject_area) %>% 
  summarise(journals_n = n_distinct(issn),
            journals_frac = sum(fractions)) %>% 
  arrange(desc(journals_frac))
# Number of articles (full and fractional counting)
subject_area_all_article <- subject_hybrid %>% 
  group_by(doi, subject_area) %>%
  summarise(n = n()) %>% 
  mutate(fractions = n / sum(n)) %>% 
  ungroup() %>% 
  group_by(subject_area) %>% 
  summarise(articles_n = n_distinct(doi),
            articles_frac = sum(fractions)) %>% 
  arrange(desc(articles_frac))
```

Next, we present disciplinary variations in hybrid OA. Table \ref{tab:subject_area_table} presents the high-level findings by AJSC subject area. In the five-year period we studied, the physical sciences (`r filter(subject_area_all_jn, subject_area == "Physical Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Physical Sciences") %>% pull(articles_n)` OA articles) and health sciences (`r filter(subject_area_all_jn, subject_area == "Health Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Health Sciences") %>% pull(articles_n)` OA articles) were the most prevalent in terms of hybrid journals with at least one OA article.  In terms of article volume, most OA articles were published in hybrid journals belonging to the life sciences (`r filter(subject_area_all_jn, subject_area == "Life Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Life Sciences") %>% pull(articles_n)` OA articles), while social sciences  (`r filter(subject_area_all_jn, subject_area == "Social Sciences") %>% pull(journals_n)`  hybrid journals, `r filter(subject_area_all_article, subject_area == "Social Sciences") %>% pull(articles_n)` articles) journals published the lowest number of OA articles. However, it is important to note that there is a large overlap between the journals’ subject areas. Therefore, Table  \ref{tab:subject_area_table} also assigns hybrid journals and OA articles fractionally to a subject area and compares it to the full counting, where journals assigned more than one subject in the AJSC classification were counted once for each field. For the remaining results, we only present the full counting.

```{r subject_area_table}
# join
inner_join(subject_area_all_jn, subject_area_all_article, by = "subject_area") %>%
  arrange(subject_area) %>%
  select(1, journals_n, articles_n, journals_frac, articles_frac) %>%
  filter(subject_area != "Multidisciplinary", !is.na(subject_area)) %>% 
  mutate(subject_area = stringr::str_to_sentence(subject_area)) %>%
  kbl(booktabs = TRUE, col.names = c("Subject area", "Journals", "OA articles", "Journals", "OA articles"),
      caption = "Subject area overview using full and fractional counting. Table reports the number of Elsevier hybrid journals with at least one open access (OA) article, and OA articles published from 2015-2019. One multidisciplinary with 61 OA articles and 11 journals with 37 OA articles that  could not be matched to the Scopus Source Title list used to obtain ASJC codes are not included.",
      format.args = list(big.mark = ",")) %>%
  add_header_above(c(" " = 1, "Full counting" = 2, "Fractional counting" = 2)) %>%
  kable_styling() 
```

```{r subject_ind}
# see 008_subject_exploration
hybrid_subject <- readr::read_csv(here::here("data", "jn_scopus_ind_subjects.csv"))
```

```{r}
subject_all_years <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(oa),
            n = sum(articles)) %>%
  ungroup() %>%
  group_by(description) %>%
  filter(subject_area != "Multidisciplinary")
# distinct journals per subject
journals_per_subject <- subject_all_years %>%
  summarise(n_journals = n_distinct(issn)) %>%
  mutate(field_label = glue::glue("{description} (n={n_journals})"))

median_all <- subject_all_years %>% 
  ungroup() %>% 
  distinct(issn, .keep_all = TRUE) %>% 
  mutate(prop = n_oa /n) %>% 
  .$prop %>% 
  median()
# rank
oa_per_subject_rank <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(oa),
            n = sum(articles)) %>%
  mutate(prop = n_oa / n) %>%
  ungroup() %>% 
  group_by(description) %>%
  summarise(median_prop = median(prop)) %>%
  arrange(desc(median_prop))
```

Figure \ref{fig:oa_sub_uptake} presents the journals' five-year OA uptake grouped by subject area and field and in comparison to the overall median (Mdn= `r median_all * 100`; excluding one journal coded as multidisciplinary). The box plot reveals some variation between the subject areas. Notably, with the exception of environmental science and earth and planetary sciences, journals from the physical sciences show a lower median proportion of hybrid OA articles, whereas most life sciences and social sciences journals ranked above average. Figure X also shows large variations among the 26 subject fields, with the highest OA rates in immunology and microbiology (Mdn=`r filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop)*100`), environmental science (Mdn=`r filter(oa_per_subject_rank, description == "Environmental Science") %>% pull(median_prop)*100`), and neuroscience (Mdn=`r filter(oa_per_subject_rank, description == "Neuroscience") %>% pull(median_prop)*100`), whilst  chemistry (Mdn=`r filter(oa_per_subject_rank, description == "Chemistry") %>% pull(median_prop)*100`), materials science (Mdn=`r filter(oa_per_subject_rank, description == "Materials Science") %>% pull(median_prop)*100`), and dentistry (Mdn=`r filter(oa_per_subject_rank, description == "Dentistry") %>% pull(median_prop)*100`) recorded the lowest OA rate.

```{r oa_sub_uptake, fig.cap='Open access (OA) uptake in Elsevier hybrid journals in percent grouped by subject area and field between 2015-19. Visualized as a box-and-whiskers plot limited to an OA share of 20\\%; the red line represents the overall median (2.5), jittered dots represent individual journals. Subject fields are ordered based on the median OA uptake.', fig.asp=1.3, out.width="90%"}
# with labels
subject_all_years_ <- 
  inner_join(subject_all_years, journals_per_subject, by = "description")

p_subject <- ggplot(subject_all_years_, aes(fct_reorder(field_label, n_oa / n, .fun = median), n_oa / n)) + 
  geom_boxplot(aes(color = subject_area)) +
  geom_jitter(alpha = .05, aes(color = subject_area), size = 1) +
  facet_wrap(~subject_area, scales = "free_y", nrow = 4) +
  geom_hline(yintercept = median_all,  colour = "red", linetype ="dashed", size = 1) +
  scale_colour_manual(NULL, values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701"),
                      guide=FALSE) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 5L),
    expand = expansion(mult = c(0, 0.05))
  ) +
  coord_flip(ylim=c(0,0.2)) +
  labs(x = NULL, y = "Five-Year Hybrid OA Uptake Rate\nper Journal") +
  theme_minimal_vgrid(font_size = 10) +
  theme(strip.text = element_text(face="bold"))
p_subject
```

### Citation Impact{.unnumbered}

```{r}
hybrid_subject_19 <- hybrid_subject %>% 
  mutate(prop = oa /articles) %>%
  filter(issued_year == "2019", !is.na(percentile)
         , prop != 0) %>%
  distinct(issn, prop, percentile, snip, sjr)

cor_test_res <- cor.test(hybrid_subject_19$prop, hybrid_subject_19$percentile, method = "spearman") 
```

Furthermore, we were interested in the relationship between the journals' field-specific impact and their OA uptake. Spearman's rho correlation coefficient was used to assess the relationship between the 2019 Source Normalized Impact per Paper (SNIP) value calculated by Scopus and the journals' OA uptake that year. Considering only journals with at least one OA article, we found a weak but positive correlation between journal impact and OA uptake ($r_s$=`r as.character(round(cor_test_res$estimate,4))`, $p<0.001$), suggesting that the relationship is not very strong, based on our data (see Figure \ref{fig:sniptest}).

```{r sniptest, fig.cap="Source Normalized Impact per Paper (SNIP) versus open access (OA) uptake 2019. Each dot represents a hybrid journal with at least one published OA article in 2019. There is only a weak, positive correlation between the two variables ($r_s$=0.1679, $p<0.001$)"}
ggplot(hybrid_subject_19, aes(prop, snip)) + geom_point(alpha = 0.2) + 
  geom_smooth(method = "loess", se = TRUE) +
  coord_cartesian(xlim = c(0., 0.7), ylim = c(0.5, 6)) +
  labs(x = "Hybrid OA Uptake", y = "SNIP") +
  theme_cowplot()
```

## Invoicing for Hybrid Open Access{.unnumbered}

### Invoice Channels{.unnumbered}

```{r}
# load eligible oa articles
hybrid_volume <- readr::read_csv(here::here("data", "hybrid_articles.csv")) 
# add open apc info
hybrid_volume <- hybrid_volume %>%
      mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
# aggregate data
sponsored_articles <- hybrid_volume %>%
  mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND")) %>%
  count(issued_year, fct_source, cc_norm)
all_articles <- hybrid_volume %>%
  group_by(issued_year) %>%
  summarise(n = n())
sponsored_art <- sponsored_articles %>% 
  group_by(fct_source) %>% 
  summarise(nn = sum(n)) %>% 
  #mutate(cumsum = cumsum(nn)) %>% 
  mutate(prop = nn / sum(nn) * 100)  %>%
 # mutate(cumprop = cumsum(prop)) %>%
  adorn_totals("row")
```

As can be seen from Table \ref{tab:invoicing_overview}, hybrid OA APC were most often invoiced to authors (n=`r filter(sponsored_art, fct_source == "Author") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Author") %>% pull(prop)`%) and to a lesser extent as part of agreements (n=`r filter(sponsored_art, fct_source == "Agreement") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Agreement") %>% pull(prop)`%). Interestingly, we also found a small number of cases where hybrid APCs were waived (n=`r filter(sponsored_art, fct_source == "Fee waived") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Fee waived") %>% pull(prop)`%).

```{r invoicing_overview}
sponsored_art %>% 
  kbl(booktabs = TRUE, col.names = c("Invoicing channel", "Hybrid OA articles (n)", "Percentage"),
      caption = "Invoice channels for hybrid OA in Elsevier journals between 2015-19.", digits = 1,
      format.args = list(big.mark = ",")) %>%
  kable_styling() %>%
  row_spec(4, hline_after = TRUE)
```

```{r els_waived}
els_waived_df <- hybrid_volume %>% 
  filter(oa_sponsor_type == "ElsevierWaived") 
sponsored_articles_prop <- sponsored_articles %>%
  group_by(issued_year, fct_source) %>%
  summarise(n = sum(n)) %>%
  mutate(perc = n / sum(n) * 100)
```

Figure \ref{fig:invoiceoverview} illustrates that over the years, Elsevier has constantly invoiced hybrid OA APCs directly to authors compared to research funders or academic consortia ("Agreement"). Between 2015 and 2019, the proportion of author-invoicing increased from `r filter(sponsored_articles_prop, issued_year == 2015, fct_source == "Author") %>% pull(perc)`% to `r filter(sponsored_articles_prop, issued_year == 2019, fct_source == "Author") %>% pull(perc)`%, while the share of invoicing through agreements slightly changed from `r filter(sponsored_articles_prop, issued_year == 2015, fct_source == "Agreement") %>% pull(perc)`% to `r filter(sponsored_articles_prop, issued_year == 2019, fct_source == "Agreement") %>% pull(perc)`%. Also, the share of fee-waived articles remained relatively stable, but we found different types of waivers. Around `r filter(els_waived_df, !is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)* 100`% of fee waivers were linked to an OA sponsor, which indicates that a third party paid the APC. For instance, `r filter(els_waived_df, oa_sponsor_name == "Académie des Sciences") %>% distinct(doi) %>% nrow()` waivers were linked to the French Académie des Sciences, presumably covering OA publication in its society journals for affiliated authors. The remaining `r filter(els_waived_df, is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)*100`% of waived articles did not disclose any OA sponsor. Moreover, Figure \ref{fig:invoiceoverview} compares the invoicing channels based on open content license variants. When Elsevier invoiced authors directly, most OA articles in hybrid journals were published under a non-commercial license (n=`r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC BY-NC-ND") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC BY-NC-ND") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Author") %>% pull(n) %>% sum()* 100`%), whereas most articles billed as part of agreements were licensed under the more permissive CC BY license (n=`r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC BY") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC BY") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Agreement") %>% pull(n) %>% sum() * 100`%).

```{r invoiceoverview, fig.cap="Development of fee-based open access (OA) publishing in Elsevier hybrid journals by invoice channel Colored bars represent the share of OA articles invoiced to authors, agreements, or waived, grouped by Creative Commons license variant. Grey bars show the total number of hybrid OA articles published in Elsevier journals from 2015-2019."}
# plot
sponsored_articles %>%
  mutate(fct_source = stringr::str_to_title(fct_source)) %>%
  mutate(fct_source = ordered(fct_source, levels = c("Author", "Agreement", "Fee Waived", "Other"))) %>%
  ggplot(aes(x = gsub("20", "'", issued_year), y = n), alpha = 0.8) +
  geom_bar(
    data = all_articles,
    color = "transparent",
    stat = "identity",
    fill = "#b3b3b3a0"
  ) +
  geom_bar(aes(fill = cc_norm), color = "white", stat = "identity", alpha = 0.8, size = 0.25,
           position = position_stack(reverse = T)) +
  facet_wrap( ~ fct_source, nrow = 1) +
  scale_fill_manual(values = 
                      c(`CC BY` = "#B52141",
                        `CC BY-NC-ND` = "#0093c7"),
                    name = "CC License Variant") +
  labs(x = "Publication Year", y = "Hybrid OA Articles") + 
  scale_y_continuous( 
                     labels = scales::number_format(big.mark = ","),
                      expand = expansion(mult = c(0, 0.05))) +
  theme_minimal_hgrid() +
  theme(legend.position="top", legend.justification = "right")
```

We also observed large differences in OA invoicing among subject fields. Table \ref{tab:invocing_by_subject} shows the number of OA articles by subject field and invoice channel. For articles in nursing, decision dciences, and pharmacology, toxicology and pharmaceutics, Elsevier predominantly invoiced authors, whereas most energy and chemical engineering articles were invoiced through agreements. Likewise, the majority of articles in materials science, chemistry and physics and astronomy were not invoiced to authors but facilitated through agreements or waived. The large share of waived APCs in physics and astronomy can be attributed to a single 2015 issue of Nuclear and Particle Physics Proceedings, which offered sponsored OA.

```{r invocing_by_subject, results='asis'}
hybrid_jn_volume <- hybrid_oa_volume %>%
  group_by(issn) %>% 
  summarise(n_oa = sum(oa))

jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))

hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
funder_shares <- hybrid_df %>%
   mutate(oa_sponsor_type = ifelse(!oa_sponsor_type %in% c("FundingBody", "Author", "ElsevierWaived"), "Other", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  group_by(oa_sponsor_type, issn) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  left_join(hybrid_jn_volume, by = "issn") %>%
  mutate(prop = n_oa_type / n_oa) %>%
  left_join(jn_subjects, by = "issn") %>%
  ungroup() %>%
  group_by(top_level, oa_sponsor_type) %>%
  summarise(oa_type = sum(n_oa_type)) %>%
  mutate(prop = oa_type / sum(oa_type))
# overall props
oa_sponsor_type_all <- hybrid_df %>% 
  #exclude multidisciplinary
  filter(issn != "2095-9273") %>%
  count(oa_sponsor_type) %>% 
  mutate(prop = n / sum(n))
subject_all <- funder_shares %>%
  group_by(top_level) %>%
  summarise(oa = sum(oa_type))

funder_overview_subject <- left_join(funder_shares, subject_all, by = "top_level") %>%
  tidyr::pivot_wider(id_cols = c("top_level", "oa"), names_from = oa_sponsor_type, values_from = prop) %>%
  ungroup() %>% 
  # remove na and multidisciplinary 
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  relocate(Agreement, .after = Author) %>%
  arrange(desc(Author))

# define heat-colors
make_pal <- scales::col_numeric(
  palette  = c("white", "#fcde9c","#faa476","#f0746e","#e34f6f","#dc3977"),
  na.color = "#808080",
  domain = NULL,
  alpha = FALSE,
  reverse = FALSE
)
# prepare row-wise grouping per subject area
subject_table <- jn_subjects %>% 
  distinct(subject_area, top_level) %>%
  inner_join(funder_overview_subject, by = "top_level") %>%
  mutate_at(4:7, function(x){
    round(x*100,0)
  }) %>%
  mutate_at(4:7, function(x){
    cell_spec(x, background = make_pal(x))}) %>%
  mutate_at(3, function(x){
    cell_spec(x, background = make_pal(x))})

subject_table_ <- subject_table %>%
  group_by(subject_area) %>%
  arrange(top_level, .by_group = TRUE) %>% 
  ungroup() 
subject_table_ %>%
  select(-subject_area) %>%
            kbl(booktabs = TRUE, "latex", format.args = list(big.mark = ","), digits = 1, escape = FALSE,
                  caption = "Invoice channels by discipline 2015-2019.",
                  position = "H",
      col.names = c("Subject", "OA articles", "Author", "Agreement", "Fee waived", "Other"),
       linesep = "",
      align = "lrrrrr") %>%
  add_header_above(c(" " = 2, "Invoice channel (in %)" = 4)) %>%
  pack_rows(index = table(subject_table_$subject_area)) %>%
  kable_styling(latex_options = c("scale_down"))
```

### Invoice Recipients{.unnumbered}

```{r}
hybrid_invoice <- hybrid_volume %>%
  filter(oa_sponsor_type == "FundingBody") %>%
  left_join(curated_invoicing_df, by = c("oa_sponsor_name")) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND"))
```

```{r license_count}
sponsor_df <- hybrid_invoice %>% 
#  filter(!is.na(Country)) %>% 
  # manual changes
  mutate(Country = ifelse(grepl("EU", Country), "European Union", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf))
```

Elsevier’s data offer more insight into invoicing. Overall, we identified 62 sponsoring institutions and organizations that received invoices for hybrid publishing fees as part of an agreement with Elsevier. We found that, by a large margin, most invoices were issued to UK-based research funders and institutions (n=`r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%), followed by the Netherlands (n=`r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%) and the European Union (n=`r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%; EU). Figure \ref{fig:invoice_sponsor_country} provides a closer look at the geographical distribution by license prevalence, highlighting the dominance of CC -BY licensed articles invoiced to sponsoring institutions in the UK, the United States, and, to a much lesser extent, the Netherlands. In contrast, the majority of articles invoiced to sponsoring institutions from Norway or representing the EU were published under a non-commercial license.

```{r invoice_sponsor_country, fig.cap="Number of centrally invoiced OA articles in Elsevier hybrid journals by country of recipient. Colored bars represent the Creative Commons (CC) license variant under which the article was published between 2015 and 2019."}
sponsor_df %>%
  group_by(country_factor, cc_norm) %>% 
  summarise(n = n_distinct(doi)) %>%
  ggplot(aes(fct_rev(country_factor), n, fill = cc_norm)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE),
            color = "white", alpha = 0.8, size = 0.25) +
  coord_flip() +
  scale_fill_manual(values = 
                      c(`CC BY` = "#B52141",
                        `CC BY-NC-ND` = "#0093c7"),
                    name = "CC License Variant") +
  labs(x = "OA Sponsor by Country", y = "Hybrid OA Articles (Agreement)") +
  scale_y_continuous(labels = scales::number_format(big.mark = ","),
                      expand = expansion(mult = c(0, 0.05)),
                     breaks = seq(0,15000, 2000)) +
  theme_minimal_vgrid() +
  theme(legend.position="top", legend.justification = "right")
```

Figure \ref{fig:sponsor_sector} shows the yearly distribution of central invoiced articles by year and institutional sector. While UK-based OA invoices were mainly addressed to discipline-specific governmental and non-profit research funders, invoices to the Netherlands and Norway were issued to national academic consortia representing the higher education sector. In 2019, Elsevier also launched similar agreements in countries with lower publication output including Hungary and Poland. Besides, we found that invoiced funding bodies from the UK and the US mainly represented discipline-specific funders, while the invoiced funding bodies from other countries were focused on a broad variety of disciplines (see Table \ref{tab:subject_profile}).

```{r sponsor_sector, fig.cap=' Development of centrally invoiced OA articles in hybrid journals by country. Colored bars represent the institutional sector OA funding bodies belong to.'}
sponsor_df %>%
  filter(!is.na(Sector)) %>%
  count(issued_year, Sector, country_factor) %>%
  mutate(Sector = stringr::str_to_title(Sector)) %>% 
  mutate(sector_level = factor(Sector)) %>%
  mutate(sector_level = fct_relevel(sector_level, "Government", "Higher Education", "Private Non-Profit", "Other")) %>%
  ggplot(aes(gsub("20", "'", issued_year), n, fill = sector_level)) + 
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE),
            color = "white", alpha = 0.8, size = 0.25) +
  facet_wrap(~country_factor) +
  scale_fill_manual("Sector", values = c("#65BADA", "#00A757", "#E5BA52", "grey50")) +
  scale_y_continuous(labels = scales::number_format(big.mark = " "),
                      expand = expansion(mult = c(0, 0.05))) +
  theme_minimal_hgrid() +
  theme(legend.position="top", legend.justification = "right",
        panel.spacing = unit(2, "lines")) +
  guides(fill = guide_legend(nrow = 2)) +
  labs(x = "Publication Year", y = "Hybrid OA Articles (Agreement)")
```

```{r}
sponsor_country <- hybrid_invoice %>%
  filter(!is.na(`Primary research focus`)) %>%
  # manual changes
  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "European Union", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf)) %>%
  mutate(`Primary research focus` = ifelse(`Primary research focus` == "SSH", "SSH", stringr::str_to_sentence(`Primary research focus`))) %>%
  group_by(country_factor, `Primary research focus`) %>%
  summarise(n = n_distinct(doi)) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n), 2) * 100)
```

```{r subject_profile}
sponsor_country %>%
  ungroup() %>%
  pivot_wider(id_cols = c("country_factor"), names_from = `Primary research focus`, values_from = prop) %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  janitor::adorn_totals(c("col", "row")) %>%
  mutate_if(is.numeric, ~paste0(., "%")) %>%
  rename(`Country/Subject area` = country_factor) %>%
  kableExtra::kbl(booktabs = TRUE,
                  caption = " Relative subject profile of centrally invoiced OA funding bodies.", align = c("lrrrrrr")) %>%
    row_spec(6:7, hline_after = TRUE) %>%
   kable_styling(latex_options = c("scale_down"))
```

```{r}
# sector/discipline
hybrid_invoice %>%
  filter(!is.na(`Primary research focus`)) %>%
  count(Sector, `Primary research focus`) %>%
  mutate(prop = n / sum(n) * 100) %>% 
  pivot_wider(id_cols = c("Primary research focus"), names_from = Sector, values_from = n) %>%
  kableExtra::kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("scale_down"))

```

### Institutional ranking{.unnumbered}

```{r open_apc_check}
hybrid_invoice <- hybrid_invoice %>%
  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
```

Table \ref{tab:sponsor_top_10} presents the top ten of `r hybrid_invoice %>% filter(!is.na(oa_sponsor_cleaned)) %>% distinct(oa_sponsor_cleaned) %>% nrow()` sponsoring bodies in our sample. Together, they accounted for around 80% of centrally invoiced APCs. The table also highlights the number of distinct journals and the share of CC BY licensed articles. Notably, UK-based research funders and the US-based Bill and Melinda Gates Foundation mainly sponsored articles published under a CC BY license. In contrast, articles funded by VSNU, the European Research Council, and Norway Institutes had a much lower proportion of CC BY.

```{r, sponsor_top_10}
hybrid_invoice_tt <- hybrid_invoice %>%
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_cleaned), "Other", oa_sponsor_cleaned)) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND")) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
hybrid_invoice_agg_general <- hybrid_invoice_tt %>%
  group_by(oa_sponsor) %>%
  summarise(journals = n_distinct(issn),
            articles = n_distinct(doi)) %>%
  mutate(prop = articles / sum(articles))
hybrid_invoice_agg_cc <- hybrid_invoice_tt %>%
  ungroup() %>%
  filter(cc_norm == "CC BY") %>%
  group_by(oa_sponsor) %>%
  summarise(cc_by = n_distinct(doi))
hybrid_invoice_agg_oapc <- hybrid_invoice_tt %>%
  filter(in_open_apc == TRUE) %>%
  group_by(oa_sponsor) %>%
  summarise(in_open_apc =  n_distinct(doi))
ind_sponsor_df <-
  left_join(hybrid_invoice_agg_general, hybrid_invoice_agg_cc, by = "oa_sponsor") %>%
  left_join(hybrid_invoice_agg_oapc, by = "oa_sponsor") %>%
  mutate(
    cc_by = cc_by / articles,
    in_open_apc = ifelse(is.na(in_open_apc), 0, in_open_apc),
    in_open_apc = in_open_apc / articles
  ) %>%
  mutate(oa_sponsor = fct_reorder(oa_sponsor, articles)) %>%
  mutate(oa_sponsor = forcats::fct_relevel(oa_sponsor, "Other", after = 0L)) %>% arrange(desc(oa_sponsor))
# overall
overall_ind_sponsor_df <-
  tibble(
    oa_sponsor = "All",
    journals = length(unique(hybrid_invoice$issn)),
    articles = length(unique(hybrid_invoice$doi)),
    prop = 1,
    cc_by = filter(hybrid_invoice, cc_norm == "CC BY") %>% nrow() / length(unique(hybrid_invoice$doi)),
    in_open_apc = filter(hybrid_invoice, in_open_apc == TRUE) %>% nrow() / length(unique(hybrid_invoice$doi))
  )

ind_sponsor_df %>%
  bind_rows(overall_ind_sponsor_df) %>%
  mutate(oa_sponsor = gsub("&", "and", oa_sponsor)) %>%
  mutate_at(4:6, function(x){
    round(x*100,0)
  }) %>%
  mutate_at(5:6, function(x){
    cell_spec(x, background = make_pal(x))}) %>%
  kbl("latex", booktabs = TRUE, escape = FALSE,
      col.names = c("Hybrid OA sponsors", "Journals", "Articles", "\\%", "CC BY", "OAPC"),
      align = c("lrrrrr"),
      linesep = "",
      caption = " Top 10 Elsevier hybrid OA sponsors by centrally invoiced articles 2015-2019.  The table furthermore presents the number of Elsevier hybrid journals with at least one invoiced article and two indicators of compliance: proportion of  CC BY licensed articles  and disclosed spending through the Open APC Initiative (OAPC.") %>%
  row_spec(11:12, hline_after = TRUE) %>%
  add_header_above(c(" " = 4, "Compliance (in %)" = 2)) %>%
   kable_styling(latex_options = c("scale_down"))
```

The table also highlights the proportion of APC funding publicly disclosed through the Open APC initiative, showing higher disclosure rates for  funders with a large CC BY share. On the other hand, APCs that were invoiced as part of transformative agreements were not publicly shared.

```{r ccby_vs_openapc}
ggplot(ind_sponsor_df, aes(in_open_apc, cc_by, size = articles)) +
  geom_point(alpha = 0.5) +
  theme_minimal_grid() +
  scale_x_continuous(limits = c(0,1), labels = scales::percent_format()) +
  scale_y_continuous(limits = c(0,1), labels = scales::percent_format()) +
  labs(y = "CC BY Share", x = "Spending info in Open APC", size = "Invoiced\nOA Articles") +
  theme(legend.position="right", legend.justification = "right")
```

Figure \ref{fig:field_funder_matrix} shows the subject profile of OA sponsors based on to the journals’ subject fields of centrally invoiced articles. As expected, the figure shows that subject-specific funders, such as the British Engineering and Physical Sciences Research Council (EPSRC), tended to focus on a limited set of subject fields. In contrast, national consortia, such as VSNU and Norwegian Institutes (NO), financed OA articles  for a broader range of fields. However, within these consortia, the subject distribution also varied. VSNU OA spending, for example, had a strong focus on articles relating to medicine, environmental science, and energy, presumably reflecting disciplines with a generally strong publication output at Dutch universities. Norwegian universities mostly published hybrid OA in engineering, environmental science and agricultural and biological sciences journals.

```{r}
#short funder labels
funder_codes <- tribble(
  ~oa_sponsor, ~oa_sponsor_short,
  "Engineering and Physical Sciences Research Council", "EPSRC",
  "VSNU", "VSNU",
  "Wellcome Trust", "Wellcome",
  "European Research Council", "ERC",
  "Medical Research Council", "MRC",
  "Natural Environment Research Council", "NERC",
  "Biotechnology and Biological Sciences Research Council", "BBSRC",
  "Bill & Melinda Gates Foundation", "Gates",
  "Economic and Social Research Council", "ESRC",
  "Norway Institutes", "NO",
  "Other", "Other")
```

```{r field_funder_matrix, out.width="90%", fig.width=8, fig.height=12}
field_funder <- hybrid_df %>%
  filter(oa_sponsor_type == "FundingBody") %>%
  left_join(funder_codes, by = c("oa_sponsor_name" = "oa_sponsor")) %>%
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_short), "Other", oa_sponsor_short)) %>%
  mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  left_join(jn_subjects, by = "issn") %>%
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  group_by(oa_sponsor, top_level) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
  mutate(oa_sponsor = fct_relevel(oa_sponsor, "Other", after = Inf))
# plot
pp <- ggplot(field_funder,
       aes(
         x = factor(oa_sponsor),
         y = reorder(top_level, desc(top_level)),
         colour = prop,
         size = n_oa_type
       )) +
  geom_point() +
  scale_color_viridis_c(
    labels = scales::percent_format(accuracy = 5L),
    option = "magma",
    name = "Percentage per Funder",
    breaks =  scales::extended_breaks(),
    begin = 0.15,
    end = 0.85
  ) +
  scale_size("OA Articles") +
  scale_x_discrete(position = "top") +
  labs(x = NULL, y = NULL) +
  theme_minimal_grid(font_size = 12) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()
  ) +
  theme(legend.position = "top", legend.justification = "right") +
  theme(legend.direction = "horizontal", 
        legend.box = "horizontal")
ggsave(here::here("figure", "funder_subject.png"), pp, dpi = 300, width = 9.5, height = 9)
```

\begin{figure}[H]

{\centering \includegraphics[width=0.95\textwidth]{../figure/funder_subject.png}

}

\caption{Top 10 Elsevier hybrid OA institutional invoice recipients in terms of articles published by journal subject. EPSRC, Engineering and Physical Sciences Research Council; VSNU, VSNU; Wellcome, Wellcome Trust; ERC, European Research Council; ERC, European Research Council; MRC, Medical Research Council; NERC, Natural Environment Research Council; BBSRC, Biotechnology and Biological Sciences Research Council; Gates, Bill and Melinda Gates Foundation; ESRC, Economic and Social Research Council; NO, Norway Institutes. }\label{fig:field_funder_matrix}
\end{figure}

# Discussion

Over the last five years, Elsevier recorded growth in the uptake of hybrid OA for each of our three indicators: The number of hybrid OA articles published per year doubled, the number of hybrid journals with at least one OA article grew by 21%, and the share of hybrid OA articles relative to closed-accessed articles in these journals increased from 2.6% to 3.7%. As @Laakso_2016, we found only a weak relationship between citation metrics like the SNIP and hybrid OA uptake and observed differences in the uptake of hybrid OA among journal subject areas. In particular, we found the highest count of hybrid OA articles in physical sciences journals (see Table \ref{tab:subject_area_table}). This was followed by the life sciences and health sciences, whereas the social sciences had the lowest count of hybrid articles. This order mostly reflects the disciplines' overall publication output. According to the Open Science Monitor by the European Commission [@OS_Monitor], the physical sciences and mathematics publish the most articles, followed by the health sciences, life sciences, and social sciences and humanities.

Such disciplinary differences in hybrid OA prevalence become more meaningful when considering the relative share of OA articles to closed-access articles. In line with previous research, we found that Elsevier journals from the life sciences and social sciences [@Jubb_2017; @Kramer_2018; @Laakso_2016] recorded greater than typical hybrid OA uptake, whereas physical sciences journals generally had a lower than typical uptake [@Kramer_2018; @Mart_n_Mart_n_2018; @Laakso_2016]. In a systematic review of OA publishing patterns, @Severin_2020 linked such disciplinary differences to socio-cultural and technological factors that shape publishing cultures and practices. Among other aspects, the authors pointed to the availability of prestigious journals, the possibility to provide OA through alternative routes, and ease of access to funding sources to cover publishing fees. As such, looking at hybrid OA uptake in isolation can only provide limited insight, so that placing it within the broader context of disciplinary publishing and OA practices is necessary. For instance, since many branches of the physical sciences have established self-archiving practices and the technical infrastructure to support these (e.g., the arXiv; @Severin_2020, @Bj_rk_2010, @Laakso_2016), researchers can provide OA through repositories (i.e., green OA), and indeed, according to the Open Science Monitor the physical sciences provide OA primarily through the green route (31.2% total OA share, 24% OA if only considering green OA). Hence, the low share of hybrid OA in our sample might be because physical scientists perceive much less of a need for hybrid OA. Further, @Severin_2020 note project-based funding structures enable OA publishing in fully OA and hybrid journals as they allow for easy integration of publishing costs. Since these funding structures are typical for the life sciences (ibid.), we would expect higher rates of OA. Data from the Open Science Monitor seems to support this notion as the life sciences record particularly high shares of full OA articles and comparatively high  shares of hybrid OA in some fields, such as biology and agricultural science. While we expected the observed hybrid OA shares for the life sciences and physical sciences, we were surprised by the high share of hybrid OA belonging to the AJSC subject area social sciences, which also includes the humanities, since articles "social scientists have reported to face significant difficulties in access to grant funding for [publication charges], as most research in the social sciences is not done by means of project-specific funding" @Severin_2020 [p. 15]. However, previous research has found comparatively low awareness and uptake of green OA among researchers in the  field [@Bj_rk_2010; @Creaser_2010], which might encourage hybrid OA uptake. Another possible explanation,is that Elsevier has a number of high prestigious social science and humanities journals [@Larivi_re_2015], which means that our findings are not representative of the general population of SSH journals. Likewise, this limitation could be extended to our findings in general that whilst not representative of the population of hybrid OA publishing, they provide valuable insight into a key player of scholarly publishing.

Finally, OA policies in favor of hybrid OA in combination with invoicing agreements with Elsevier might be a driver. For instance, we found that nearly half of articles in arts and humanities were facilitated through agreements, which was above-average. This is in line with previous research, suggesting that OA policies and funder mandates have a stronger effect on OA uptake than disciplinary publication cultures [@Huang_2020; @Larivi_re_2018].  While these aspects are important drivers of both full and hybrid OA, faculty members generally prioritize journal prestige and the journal impact factor [@Niles_2020] in publishing decisions, which, in combination with the availability of OA funding, might encourage publishing hybrid OA [@van_der_graaf_maurits_2017].

Our overall investigation of invoice channels and recipients of hybrid OA APCs found that the majority of Elsevier hybrid articles were invoiced to the author (n=41,725; 58.2%). However, it is important to note that although Elsevier's metadata classifies the sponsor type as "author", this does not necessarily mean that authors paid for APCs out of their own pocket but rather just that the APC was not invoiced through an agreement. Hence, it is possible that APCs invoiced to the author were covered by institutional funds or as part of research grants. For instance, a recent Springer Nature survey found that hybrid OA was predominantly supported through institutional and funder sources (71%), followed by OA agreements (34%), while only 6% were paid from personal funds or savings [@Monaghan_2020]. We also observed notable differences in licensing among invoice types. The majority of hybrid OA articles, regardless of the invoice type, were licensed under the more restrictive CC BY-NC-ND license. There is a lack of dedicated studies on licence selection, but indications so far suggest that when given a choice authors tend to select more restrictive variants for their published works [@Fraser_2020; @Rowley_2017; @Van_Noorden_2013]. On the other hand, when hybrid articles were invoiced through agreements, most articles were licensed under the more liberal CC BY license. As several funding bodies mandate CC BY licenses, including the UK research funders that account for 62% of our "agreement" subsample, this result is perhaps not surprising but suggests the effectiveness of such agreements.

Based on our findings, around a third of the articles were invoiced directly through OA agreements (e.g., research funders, national library consortia). Strikingly, sponsors based in the UK accounted for more than half of these , and a quarter was charged to sponsors from other European countries. This finding reiterates reports from previous studies that the UK's OA profile differs markedly from other countries (cf. @Jubb_2015; @Jubb_2017) and points to the effects of science policy. To increase OA to research publications, the UK implemented centralized APC funding and embraced hybrid OA as a transition model. Within five years of the publication of the Finch Report in 2012, the UK recorded an 18% increase in immediate OA, coupled with a rise in hybrid OA from 2.7% in 2012 to 15.4% of all articles in 2016 [@Jubb_2017], which can be attributed to the availability of APC funding from Research Councils UK (RCUK). Indeed, RCUK block grants have been the largest single source of OA funds in the UK—the Wellcome Trust, and more recently transformative agreements [@Tickel_2018; @Jubb_2017]. However, since then publishing expenditures of UK universities have been rising rapidly (ref), so that several universities stopped supporting hybrid OA in late 2018-2019 [@oxford_2019; @walker_2019; @Castro_2019; @birmingham]. This development might explain the slight decrease in hybrid articles invoiced to the UK our data showed around that time (see Figure \ref{fig:invoice_sponsor_country}).

An essential part in the absence of transparency in scholarly publishing and especially in hybrid OA has been the lack of transparency in publishers’ article management. This has prohibited consumers to get a comprehensive and comparative overview of their publishing activities and the associated costs and, thus, limited their negotiating power. However, certain initiatives in Europe like the ESAC Initiative have begun to press for improved and more transparent workflows for transformative agreements [@Geschuhn_2017]. In particular, the ESAC Workflow Recommendations seek to increase efficiency and transparency in the publication workflow through improved invoicing and reporting processes and metadata delivery, including article-based machine readable information about the OA sponsor. Through this study we demonstrated the utility and benefits of such  publisher-provided metadata about hybrid OA funding and highlighted the need for extending these to include comprehensive information about the licensing and the particular OA type.

Usually identifying delayed OA articles and distinguishing them from other types of OA is challenging since there is no publicly available list of such journals. For instance, Taylor & Francis’ (2021) website indicates that some of their journals participate in a delayed or promotional OA program but no further information is provided as to which journals these are. However, Elsevier provides a list of journals enrolled in the "Open Archive" programme and further tags delayed OA articles in the machine-readable XML version of the article. This enables easy identification and more accurate assessments of hybrid OA, which allows to take into account the substantial amount of delayed OA content in future hybrid OA studies using Elsevier data.

While this study advances our knowledge about hybrid OA uptake and invoicing, the limitations leave room for future research. This study focused on only one publisher, which limits the generalizability of our findings since Elsevier’s attributes such as composition of the journal portfolio, mix of business models, pricing, and promotion of various options cannot be assumed to be representative of scholarly journal publishing in general. Though the invoicing data extracted from the XML files enables unprecedented accuracy and comprehensiveness in the study of hybrid OA, it likely does not disclose the actual OA sponsor in all cases. Most articles were invoiced to authors, but it remains unclear how often authors themselves pay for APCs or invoices are covered through institutional OA funds or research grants. Further research into this topic would increase our understanding of OA funding streams outside of publishing agreements. Further, our study demonstrates that to achieve comprehensive mapping and understanding of the financial flows involved in OA funding, complex, in-depth country-specific meso-level analyses is needed. Such studies would ideally draw on various data sources, such as publicly available bibliometrics and metadata, information about significant national research funders and their OA policies, information about the national library and possible consortia landscape including their publishing agreements, which might require Freedom of Information Requests.

This study provides a snapshot of hybrid OA for the largest journal publisher prior to the impact of the implementation of Plan S requirements. While many Plan S signatories have already had strong OA policies, this harmonized approach is likely to affect publishing decisions of funded authors,the licencing of their articles, and the offerings and pricing of publishers. Since the new requirements are applicable to research funded from 2021 onwards, a comparative study focused on articles  invoiced to Plan S signatories would be a fruitful endeavor.

In a similar vein, future studies could  combine invoice information with research funding information since these contribute different pieces to the funding stream puzzle. This could also provide more insight into the actual APC funding streams behind the author-invoiced articles.

# Conclusion

The primary aim of this empirical study was to investigate Elsevier’s hybrid OA publishing from 2015-2019 to better understand the volume and invoicing of hybrid OA and present a novel, data-driven approach for such analyses. Our results indicate that although the number of hybrid OA articles has increased over time, hybrid OA uptake has remained low. Notably, hybrid APCs were most often invoiced directly to the authors or through institutional agreements, where only a few funding bodies were the main driver of hybrid OA. Finally, our findings highlight that publisher-provided metadata about the invoicing channels  of (hybrid) OA can facilitate research into and increase the understanding of the financial flows of OA publishing.

Ever since its inception, hybrid OA has been a challenging subject to study due to the lack of standardised ways for publishers to flag such content and APC funding data being limited to studies based on surveys and articles’ acknowledgements sections. This study presented a novel approach to studying OA APC invoicing  that is based on publicly available publisher-provided metadata, which can be used on on its own or in combination with other public data sources to gain more detailed and comprehensive insights into hybrid OA uptake and invoicing than ever before. If more publishers reported OA invoicing data on the article level and in a machine-readable format, this would increase transparency and enable  enhanced monitoring of the scholarly journal landscape over time. As hybrid OA has increasingly become a central element of OA policies of research funders and libraries,  customer organisations could require that invoicing information is added to the article-level metadata. As long as publishers do not provide this data in a structured and comprehensive format, they prevent consumers from benchmarking prices and therefore hinder price competition as the only alternative to gain comparative insight is self-reported APC expenditures that research funders and libraries voluntarily report to the Open APC initiative. The Open APC initiative is a valuable service but it cannot replace publisher-provided data since global participation is limited.

From recent science policy developments in Europe it appears that Big Deals have gained support and remain firmly in place in the form of transformative agreements. However, in addition to research funders and libraries that increasingly make publishing contracts public, it is on the publisher to provide  metadata as presented here to improve the transparency in scholarly publishing by linking invoicing data to bibliometrics. From this study we can affirm that hybrid OA is complex as the financial flows come from various stakeholders and invoicing forms, it is not just libraries or consortia, but authors and funders are also directly involved. As of right now, providing  machine-readable article-level invoicing data is not common amongst publishers, which limits the potential for broader studies.

# Data Availability

All data and code used for this study is openly available on GitHub, including version history:

<https://github.com/njahn82/elsevier_hybrid_invoicing>

Formal analyis is written in R Markdown and thus replicable.

# References