### 2021-05-19 analysis/manuscript.Rmd
align text

---
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: apa6.sty
csl: apa.csl
bibliography: references.bib
---

```{r knitr_options, , echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300,
  fig.pos = "H", 
  out.extra = "",
  dev= c("png", "cairo_ps")
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(formatC(x, big.mark = ",", format = "f", digits = 1, drop0trailing = TRUE))
    } else{
      return(x)
    }
  }
)
```

```{r libraries}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(gt)
library(janitor)
library(kableExtra)
library(stringr)
```
# Introduction

The rise of open access (OA) has added complexity to
scholarly publishing, particularly concerning transparency of economic
dimensions. Financial transparency in journal publishing has long been
lacking because Big Deal subscription contracts between academic
institutions and large academic publishers usually include
confidentiality clauses that prohibit publicly sharing
the scope or pricing of such agreements [@Frazier_2001; @Bergstrom_2014; @Larivi_re_2015]. Further, the distributed payment processes
to access and publish scholarly articles add to the opacity of
the financial flows of scholarly publishing [@Lawson_2016] as they can
involve several actors simultaneously, including authors, research
institutions, libraries, and funders.

The problem with this lack of transparency becomes more apparent in the
case of hybrid OA. In this OA business model, individual articles can be
made openly available by paying an article processing charge
(APC), while the journal as a whole remains subscription-access. Hence,
hybrid OA was originally introduced as a low-risk transitional model
that allows journals to gradually convert to full OA and reduce
subscription costs as the OA uptake increases [@Prosser_2003]. Many
large subscription publishers have since incorporated hybrid OA, but there have been concerns that the hybrid model allows for double dipping---receiving two payments for one article, the APC and
subscription fees [@Shieber_2009; @Prosser_2015]. While publishers
assure they adjust their pricing
[@Mittermaier_2015], without transparency around the uptake of hybrid OA
and both revenue streams, such claims are impossible to evaluate. This is especially concerning because hybrid OA has recently gained popularity after a decade of only slow uptake [@Bj_rk_2012]. The main drivers of this development have been research institutions and funders that implemented  OA policies, allocated OA funds, and established
formal agreements with publishers that allow
affiliated authors to publish OA free-of-charge [@Laakso_2016].

The recent introduction of transformative agreements, an evolving
concept describing contracts that shift library spending from subscriptions to OA
[@Borrego_2020; @Hinchliffe_2019], might perpetuate the lack of transparency because pricing is often based on traditional subscription costs. Hence, the demand for
publisher-provided data has increased. More
transparency about hybrid OA uptake and funding could  facilitate the assessment and adjustment of publisher contracts [@Schimmer_2015], enhance OA mandate compliance monitoring, and avoid double-dipping [@Larivi_re_2018]. However, previous studies have noted an absence of transparency in hybrid OA publishing and a considerable lack of
publicly available and standardized data [@Laakso_2016; @Lawson_2015; @Pinfield_2016].

This paper investigates to what extent publisher-provided
data can increase the transparency around hybrid OA by taking Elsevier,
the largest scholarly journal publisher [@Larivi_re_2015], as the starting point for an empirical
analysis of hybrid OA uptake and invoicing. Our approach is based on
openly available data sources and combines scholarly metadata
from Crossref, a DOI registration agency, with
 APC invoicing data from Elsevier. The
compiled data enable us to analyze the 
hybrid OA uptake in Elsevier journals between  2015 and 2019 by license and subject, and to compare it with
Elsevier's full and delayed OA program. Further, we determine whether
hybrid APCs were waived or charged to authors or as part of publishing agreements. In the latter case, we 
examine the invoiced academic consortia and research funders. As such, our findings have implications for research and OA
policy and highlight the potential of publisher-provided data for
large-scale, instantaneous analysis of hybrid OA uptake and invoicing.

# Background

## Prevalence / uptake studies

Since there is no standardized way to measure the uptake of hybrid OA, estimates can vary greatly depending on how uptake is operationalized--for example, through the number of hybrid journals with at least one OA article, the average number of OA articles per hybrid journal, or the share of OA articles compared to subscription articles.

In the first article-level study on hybrid OA, @Laakso_2016
reported that between 2007 and 2013, the five largest
publishers--Elsevier, Sage, Springer, Taylor \& Francis, and
Wiley--recorded growing numbers of hybrid journals that coincided with a
twenty-fold increase in hybrid OA (from 666 to 13,994). As a
result, hybrid journals with at least one OA article more than doubled
in number (1,082 in 2009, 2,714 in 2013) but fell by 31\% relative to
the total number of hybrid journals. Laakso and Björk's
(2016) exploratory approach yielded a comprehensive overview tracking the growth of hybrid
OA over time and across multiple publishers, but due to the amount of
manual data collection and cleaning it
is unsuitable for repeated use. Minimizing such manual tasks,
Nelson \& Eggett (2017) narrowed their focus to one publisher--the
American Chemical Society (ACS)--and requested a list of all hybrid OA
articles published from 2006-2011 from the ACS. To assess the
hybrid OA uptake, the authors aggregated and compared the number of OA
(n=814) and subscription articles (n=27,621), finding that over the five
years, 2.9\% of ACS articles were published as hybrid OA.

Contrary to the previous two studies, @Kirkman_2018 took a research
funder as a starting point, reporting a 26\% share of hybrid OA for
articles funded by the Australian National Health and Medical Research
Council from 2012-2014 (816 of 3,190). Kirkman collected research
articles using the Funding Agency and Funding Text search fields on Web
of Science, and classified articles as OA when they were freely
available on the publisher website, and further distinguished hybrid OA
articles based on journal-level information from the publisher website,
such as hybrid OA journal and APC lists. Another approach is presented
by @P_l_nen_2020, who used current research information system
(CRIS) data---institutional publication data---to study the extent of OA
among Finnish research publications from 2016-2017 (n=34,507 research
articles). As part of the analysis, the authors identified hybrid OA
articles through a dedicated OA status metadata field that has been
mandatory to report in Finland since 2016 and found a 7\% share of
hybrid OA. However, the drawbacks of Kirkman's (2018) and Pölönen et
al.'s (2020) approaches are that such data sources might not be openly
available or, regarding CRIS data, might not be available at all because
not all countries maintain comprehensive institutional publication data.

Other studies assessed the share of (hybrid) OA for the entire corpus of research articles. Using DOI-assigned research articles indexed in the Web of Science Core Collection
 as a benchmark, @Mart_n_Mart_n_2018 investigated how
many research articles from 2009 and 2014 (n=2,610,305 in total) are freely available through
Google Scholar. After identifying OA articles through licensing
information from Crossref, the authors estimated the share of hybrid OA
at around 0.5\% in 2009 and 1.5\% in 2014. Similarly, in a study assessing the OA level among all scholarly articles and that experienced by Unpaywall users, Piwowar et
al. (2018) obtained license information through Crossref and web-scraping. The study found 3.6\% of 100,000 random Crossref DOIs were hybrid OA
and pointed towards a growing trend in recent years that recorded a
hybrid OA share of 9.4\% of all articles published in 2015.

Since its launch, Unpaywall has become widely used in bibliometric research and rankings [@Robinson_Garcia_2020; @Huang_2020]. However, the lack of standardized and comprehensive publisher-provided data has required several updates to Unpaywall to improve hybrid OA identification and differentiation from  delayed OA [@Piwowar_2019; @Unpaywall_oa_license; @Unpaywall_types], illustrating ongoing challenges in tracking and comparing hybrid OA prevalence over time. After changes in Unpaywall's approach to detect hybrid OA, Piwowar et al. (2019) estimated a hybrid OA share of 4\% amongst all articles published by July 2019.

## Financial studies

Pinfield et al.'s (2016) analysis of APC payment records provided by 23
United Kingdom (UK) higher education institutions revealed a sharp increase in central
payments from 2007-2013, which was largely attributed to the
introduction of block grants by Research Councils UK (RCUK) and
non-compliance sanctions by the Wellcome Trust. Moreover, the study
showed that OA fees were paid almost exclusively through block grants (92\%), and only a small number of APCs were paid
through internal funding (7\%). In contrast, a recent Springer Nature
survey found that authors draw on a range of funding sources to cover OA
fees, such as dedicated institutional OA funds, block grants, OA
agreements, or research grants (Monaghan et al., 2020). Most hybrid OA
authors were supported through dedicated institutional OA funds (43\%,
excluding block grants) and OA agreements with Springer Nature (34\%).
The differences between these two studies might be due to Monaghan et al.'s
(2020) focus on Springer Nature articles or due to} different policy and funding arrangements---considering the introduction of OA
agreements since  @Pinfield_2016 and Monaghan et al.'s
(2020) more regionally diverse sample.

Regional and policy differences in APC payments also came to light in a
study by @Jahn_2016 that analyzed APC records from 30 German
higher education and research institutions, the Austrian Science Fund (FWF), Jisc, and the Wellcome
Trust. In particular, the study revealed large differences in the amount
of hybrid OA funded from 2014-2015. Whilst hybrid OA accounted for less
than 1\% of APCs paid by German institutions (23 of 3,846), the three
non-German research funders recorded a hybrid share of 75\% (11,533 of
15,779). According to @Jahn_2016, this could point towards
differences in science policy, such as hybrid OA being supported by the
three non-German research funders but not by Germany's largest national
funder, the German Research Foundation (DFG). Another possibility is that German hybrid OA fees were
paid from budgets not reported to the Open APC initiative, a crowd-sourcing effort [@Pieper_2018] from where
the authors acquired data. Among these unreported funds
are research grants and research unit budgets, which author surveys
identified as APC funding sources [@Monaghan_2020; @Graaf_2017]. As such, Jahn and Tullney's (2016) findings could reflect
the complexities and potential limitations of institutional OA spending
data that @Pinfield_2016 and @Monaghan_2020 attributed
to incomplete or missing records.

In recent years, national consortia in Europe have negotiated publishing agreements covering hybrid OA fees for affiliated authors (Borrego et al., 2020). While these have improved invoicing workflows, internal assessments of hybrid OA uptake and invoicing remain challenging because transparent and comparative data have largely remained absent [@Marques_2020]. Such publicly available information might be scarce because publishers and institutions lack or withhold such data, or due to confidentiality clauses [@Marques_2020; @Monaghan_2020]. Several European consortia have collaborated through the Efficiency and Standards for Article Charges (ESAC) Initiative and Knowledge Exchange to standardize metadata requirements across publishers as great variation has been found between different countries and publishers [@Marques_2019].

## Research questions and aims

In this paper, we focus on Elsevier, a prominent example in recent hybrid OA uptake and financial studies [@Pinfield_2016; @@Laakso_2016]. Elsevier's OA portfolio presents the challenges in examining hybrid OA described above. For instance, distinguishing between different OA types. Elsevier supports delayed [@Els_Archive], hybrid, and full OA, including so-called mirror journals---full OA counterparts of hybrid journals addressing OA policies opposed to hybrid OA [@Harrison_2019]. Further, Elsevier processes APC invoices through various channels, such as agreements with research funders and library consortia or, in the absence thereof, the authors (Elsevier,
n.d.-a, n.d.-c). Surprisingly, we found Elsevier article-level metadata embedded in XML full-texts indicating the articles' OA status and invoicing.  Leveraging this publicly available data, we 
address the lack of transparency around hybrid OA noted in previous studies.
In particular, we use this novel approach to answer the following
questions:

* What was the uptake of Elsevier's hybrid OA publishing option between 2015 and 2019?
* Through which channels were hybrid APCs invoiced, and who were the recipients?

# Methodology

For this study, we collected data relating to Elsevier's hybrid OA
option by drawing on multiple freely available data sources. We
identified Elsevier hybrid journals through Elsevier's APC list and
supplemented our sample with Crossref metadata and text-mined invoicing
data to investigate the invoicing of immediate OA articles provided
under a Creative Commons (CC) license in subscription-based journals. Figure
\ref{fig:workflow} visualizes the automated workflow we used to collect
data from Elsevier and Crossref.

\begin{figure}[H]

{\centering \includegraphics[width=0.95\textwidth]{../figure/els_flow.pdf}

}

\caption{Data Collection Workflow to Obtain Article-Level OA and Invoicing Data.
}\label{fig:workflow}
\end{figure}

First, we identified hybrid journals through an Elsevier APC price list
from May 2020 provided by @Matthias_2020. We excluded journals that transitioned from hybrid to full OA and
reverse-flip journals that flipped 
from full to hybrid OA [@Matthias_2019]. Overall, we identified 1,970 unique hybrid journals that published between 2015 and 2019.

Next, we used an openly available Crossref database snapshot [@Crossref_2020], which contains all Crossref records registered until March 2020,
to calculate the journals' combined article volume for the five-year
period from 2015-2019. We only included articles published in regular
issues aside from supplements containing conference contributions like
meeting abstracts, indicated by non-numeric pagination. We excluded
non-scholarly journal content, such as the table of contents, following Unpaywall's paratext recognition approach
 [@Unpaywall_para], which we expanded to include patterns indicating
corrections. Furthermore, we categorized the articles by subject
according to the All Science Journal Classification code (ASJC) and
added citation indicators from the Scopus Source Title list (version
June 2020) based on the journal they were published in.

We identified OA articles
through CC licenses in Crossref metadata records (Hendricks et al., 2020) and then downloaded the XML
version of all CC-licensed articles published in a hybrid journal. From
the XML files, we obtained the articles' OA status and invoicing information (see Table \ref{tab:els_xml}). We determined whether articles were immediate or delayed
OA using the XML node \texttt{openArchiveArticle}, which indicates if an article was made freely available under Elsevier's Open Archive program, and measured the
uptake of hybrid OA. Moreover, we obtained the invoice channels and recipients of hybrid OA APCs. Using the \texttt{openaccessSponsorType} node, we distinguished between four invoice channels, including
invoices billed to authors, as part of publishing
agreements with funding bodies (cf. @Els_Agreements), exempted through fee waivers (e.g., in ``cases of genuine need'' or due to
society or university sponsorships, cf. @Els_Pricing), and other types not
specified by Elsevier. Finally, when hybrid APCs were invoiced as part
of agreements, we identified invoice recipients through the
\texttt{openaccessSponsorName} node. It is worth noting that while this article focuses on OA articles published in hybrid journals, the same data is also available for articles published in full OA journals. Elsevier does not provide the invoiced APC in either case.

\begin{table}[H]
\caption{\label{tab:els_xml}Metadata in Elsevier XML Full-Texts.}
\centering
\begin{tabular}[t]{lp{10cm}}
\toprule
XML node & Description\\
\midrule
\texttt{openaccessArticle} & Was the article open
access? \\
\texttt{openaccessSponsorType} & Invoice channel.\\
\texttt{openaccessSponsorName} & Invoice recipient.\\
\texttt{openArchiveArticle} & Was open access provided through
Elsevier's Open Archive program?\tabularnewline \\
\bottomrule
\end{tabular}
\end{table}

We manually classified invoice recipients based on their institutional
sectors, countries, and primary research areas. Following the OECD's
Frascati Manual [@Frascati, p. 91], we coded for four sectors: business
enterprise, government, higher education, and private non-profit. 
Due to the low article volume,
we combined the business enterprise sector with invoice recipients Elsevier listed as "authors" and
"third-party sponsor"
 into "Others". Moreover, we categorized invoice
recipients according to the countries representing their
scope of funding and based on the following primary research areas
health sciences, life sciences, physical sciences and mathematics,
social sciences and humanities, broad (i.e., multiple
research areas), and unknown. Further, we compared Elsevier's invoicing data
with institutional spending data from the Open APC initiative.

Throughout this mostly automated data gathering and analysis process, we
used tools from the Tidyverse [@tidyverse] for the R
programming language [@r]. To allow for efficient data
manipulation and retrieval, we imported the Crossref dump to Google
BigQuery, applying the rcrossref [@rcrossref] parsers to
extract relevant metadata fields. We used crminer
[@crminer] to obtain the XML-full texts from Elsevier.

# Results{.unnumbered}

This section first presents the results of our analysis of hybrid OA
uptake in Elsevier's journal portfolio with a view to licensing,
disciplinary differences, and citation impact. Then, we present a
descriptive analysis of Elsevier's invoicing data, highlighting
licensing and disciplinary differences for invoicing channels and invoice recipients.

```{r data}
# journal-level data
els_jns <- readr::read_csv(here::here("data", "els_historic_jns.csv"))
els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
# article-level data
hybrid_oa_volume <- readr::read_csv(here::here("data", "hybrid_oa_volume.csv"))
# invoicing analysis
curated_invoicing_df <- readr::read_csv(here::here("data", "curated_invoicing_data.csv"))
# open apc data
o_apc <- readr::read_csv(here::here("data", "o_apc.csv"))
```

## Uptake of Hybrid Open Access{.unnumbered}

### Overview{.unnumbered}

```{r}
# journals by year and oa model
els_jns_count <- els_jns %>%
  group_by(year, oa_model) %>%
  count(year) %>%
  group_by(year) %>%
  mutate(prop = n / sum(n)) %>%
  mutate(oa_model_fct = fct_rev(as_factor(oa_model)))
# overall count
 els_overall <- els_jns %>% 
  group_by(oa_model) %>% 
  summarise(n = n_distinct(issn))
```

```{r, rev_flipped}
flipped_jns <- els_jns %>%
  distinct(issn, oa_model) %>%
  group_by(issn) %>%
  filter(n() > 1)
flipped_jns_ind <- els_jns %>%
  filter(issn %in% flipped_jns$issn) %>%
  group_by(issn) %>%
  slice(which.max(year)) %>%
  ungroup() %>%
  count(oa_model)
```

```{r hybrid_oa_per_year}
# journal and article metrics **per year**
hybrid_oa_per_year <- hybrid_oa_volume %>%
  group_by(issued_year) %>%
  # distinct jns with at least one article per year
  filter(oa > 0) %>%
  summarise(
      # distinct jns with at least one article per year
    n_journals = n(),
    # article metrics
    articles_n = sum(articles),
    articles_mean = mean(articles),
    articles_sd = sd(articles),
    # oa articles metrics
    oa_n = sum(oa),
    oa_mean = mean(oa),
    oa_sd = sd(oa)
    ) %>%
  mutate(oa_uptake = oa_n / articles_n)
# oa uptake per year
oa_uptake_per_year <- hybrid_oa_volume %>% 
  filter(oa > 0) %>%
  mutate(prop = oa / articles) %>%
  group_by(issued_year) %>%
  summarise(oa_uptake_mean = mean(prop),
            oa_uptake_sd = sd(prop))
# merge indicators
ind_per_year <- inner_join(hybrid_oa_per_year, oa_uptake_per_year, by = "issued_year")

# five-years article volume
all_jn_articles <- hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    articles_p_jn = sum(articles)
    ) %>%
  ungroup() %>%
  summarise(
    n_journals = n(),
    articles_n = sum(articles_p_jn),
    articles_mean = mean(articles_p_jn),
    articles_sd = sd(articles_p_jn)
  )
# five-years OA article volume
all_oa_articles <- 
  hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    oa_p_jn = sum(oa)
    ) %>%
  ungroup() %>%
  summarise(
    oa_n = sum(oa_p_jn),
    oa_mean = mean(oa_p_jn),
    oa_sd = sd(oa_p_jn)
  )
# five-years OA uptake rate
oa_uptake <- 
  hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    articles_p_jn = sum(articles),
    oa_p_jn = sum(oa)
    ) %>%
  mutate(prop = oa_p_jn / articles_p_jn) %>%
  ungroup() %>%
  summarise(oa_uptake = sum(oa_p_jn) / sum(articles_p_jn),
            oa_uptake_mean = mean(prop),
            oa_uptake_sd = sd(prop)
  )
# bringing it all together
five_years_ind <- bind_cols(issued_year = "2015-19", all_jn_articles, all_oa_articles, oa_uptake)

ind_summary_table <- ind_per_year %>%
  mutate(issued_year = as.character(issued_year)) %>% 
  bind_rows(five_years_ind) %>%
  mutate(across(contains("uptake")) * 100) %>%
  pivot_longer(n_journals:oa_uptake_sd, names_to = "indicator", values_to = "count") %>%
  pivot_wider(names_from = issued_year, values_from = count)
```

Between 2015 and 2019, `r hybrid_oa_volume %>% distinct(issn) %>% nrow()` of `r els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` (`r  hybrid_oa_volume %>% distinct(issn) %>% nrow() / (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n))*100`%) hybrid journals published at least one OA article. Together these journals published `r five_years_ind %>% pull(oa_n)` OA articles, which represents `r five_years_ind %>% pull(oa_uptake)*100`% of their total article volume (n=`r five_years_ind %>% pull(articles_n)`). Table \ref{tab:overview} presents  findings for each year and the aggregated five-year period, illustrating moderate growth in hybrid OA over time. Each year, the number of hybrid journals
with at least one OA article increased, as did the number of immediate
OA articles in these journals, which nearly doubled from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_n)` in 2015 to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_n)` in 2019. However, since the journals' total article output also grew over time, the relative share of OA articles
in hybrid journals only increased slightly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake)*100`% to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake)*100`%.

```{r overview}
table_descriptives_ind <- tribble(
  ~ind_group, ~ind, ~` `,
    "Elsevier Hybrid Journals with ≥ 1 OA article (n)", "n_journals", "Total",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "articles_n", "Total ",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "articles_mean", "Avg.",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "articles_sd", "SD",
  "OA Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "oa_n", "Total",
  "OA Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "oa_mean", "Mean",
  "OA Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "oa_sd", "SD",
  "Share of Elsevier Hybrid Journals with ≥ 1 OA article (%)", "oa_uptake", "Total",
  "Share of Elsevier Hybrid Journals with ≥ 1 OA article (%)", "oa_uptake_mean", "Mean",
  "Share of Elsevier Hybrid Journals with ≥ 1 OA article (%)", "oa_uptake_sd", "SD"
)
inner_join(table_descriptives_ind, ind_summary_table, by = c("ind" = "indicator")) %>%
  dplyr::select(-ind, -ind_group) %>%
  kableExtra::kbl(booktabs = TRUE, format.args = list(big.mark = ","), digits = 1, escape = TRUE,
                  caption = "Elsevier Hybrid OA Uptake  2015-2019.",
                  position = "H"
) %>%
  kable_styling() %>%
  pack_rows("Elsevier Hybrid Journals with ≥ 1 OA article (n)", 1, 1) %>%
  pack_rows("Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", 2, 4) %>%
  pack_rows("OA Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", 5, 7) %>%
  pack_rows("Share of Elsevier Hybrid Journals with ≥ 1 OA article (%)", 8, 10)
```

```{r}
perc_df <- hybrid_oa_volume %>%
  mutate(prop = oa / articles) %>%
  filter(oa > 0) %>%
  group_by(issued_year) %>% 
  summarize(median = median(prop),
            p_10 = quantile(prop, 0.1),
            p_95 = quantile(prop, 0.95),
            iqr = quantile(prop, 3/4) - quantile(prop, 1/4)) %>%
  ungroup()
```

Table \ref{tab:overview} also shows that the average OA share per journal was higher than the overall percentage. To consider these variations across journals, Figure \ref{fig:boxuptake} illustrates the range of OA uptake rates as a box-and-whiskers plot by year. The bold line shows the median OA prevalence across journals, which slightly rose from `r filter(perc_df, issued_year == 2015) %>% pull(median) * 100`% in 2015 to `r filter(perc_df, issued_year == 2019) %>% pull(median) * 100`% in 2019}. Notably, the upper quartiles and
whiskers stretch farther over the years, which indicates that the range
of uptake among journals with an above-average proportion of OA articles
increased over the years. Nevertheless, the prevalence of hybrid OA
remained relatively low; in 2019, 95% of Elsevier hybrid journals published ≤`r perc_df %>% filter(issued_year == 2019) %>% pull(p_95) * 100`% of their articles OA.

```{r boxuptake, fig.cap = "Open Access Uptake per Elsevier Hybrid Journal (Per Year)."}
hybrid_oa_volume %>%
  mutate(prop = oa / articles) %>%
  filter(oa > 0) %>%
ggplot(aes(factor(issued_year), prop)) + 
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) + 
  coord_cartesian(ylim=c(0,0.2)) +
  theme_minimal_hgrid() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L),
      expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Publication Year",
       y = "Hybrid OA Uptake")
```

```{r oa_volume}
# see analysis/006_license.Rmd how the data was obtained

license_df <- readr::read_csv(here::here("data", "year_per_oa_type_and_license.csv"))
# all oa
all_oa <- license_df %>% 
  group_by(type) %>% 
  summarise(n = sum(articles)) %>%
  pull(n) %>%
  sum()
# all articles
all_els <- license_df %>%
  distinct(all_articles, issued_year) %>%
  pull(all_articles) %>%
  sum()
# summary by type
summary_oa_type <- 
  license_df %>% 
  group_by(type) %>% 
  summarise(n = sum(articles)) %>%
  mutate(prop = n / sum(n))
# mirror journals
mirror_jn <- readr::read_csv(here::here("data", "x_journals_volume.csv"))
```

During the same period, Elsevier published OA to `r all_oa` articles in total (i.e., full, hybrid, and delayed OA combined), representing `r all_oa / all_els * 100`% of the total article volume (n=`r all_els`). Looking at Figure \ref{fig:oa_volume_fig}, it is apparent that the OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`; `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged delayed OA (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`; `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) and full OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`; `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r sum(mirror_jn$all_articles)` articles published in `r length(unique(mirror_jn$container_title))` mirror journals.

```{r oa_volume_fig, fig.cap="Elsevier OA Article Volume 2015-2019 by (A) OA Type and (B) Open Content License."}
# see analysis/006_license.Rmd how the figure was created

knitr::include_graphics(here::here("figure", "license_portfolio.png"))
```

### License Prevalence{.unnumbered}

```{r cc_by_by_type}
cc_by_by_type <- license_df %>%
  group_by(type, license_code) %>%
  summarise(n = sum(articles)) %>%
  mutate(prop = n / sum(n)) %>%
  filter(license_code == "CC BY")
```

As far as we observed, OA articles in Elsevier hybrid journals were published under two possible CC licenses--CC BY or CC BY-NC-ND. CC BY allows others to distribute, remix, adapt, and build upon the licensed work, including for commercial purposes, as long as the original author is credited. In contrast, CC BY-NC-ND is less permissive as it prohibits commercial reuse and derivatives. In the five-year period from 2015-2019, the proportion of hybrid OA articles published under a CC BY-NC-ND license marginally increased and maintained a greater share than that of CC BY (see Figure \ref{fig:oa_volume_fig}-B). Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid journals had the highest number and proportion of openly available articles licensed under CC BY (n=`r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(n)`; `r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(prop) *100`%) compared to full OA journals  (n=`r filter(cc_by_by_type, type == "full_oa") %>% pull(n)`; `r filter(cc_by_by_type, type == "full_oa") %>% pull(prop) *100`%) and Elsevier's delayed OA program (n=`r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(n)`; `r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(prop) *100`%) . Most  delayed OA articles were provided under an Elsevier user license (Els-User), which prohibits reuse for commercial purposes and redistribution.

### Subject Area and Field{.unnumbered}

```{r subject_data}
# Load hybrid article-level dataset and journal classification
# see
jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))
# article level data set, see data-raw/hybrid_jns_articles.R
hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
subject_hybrid <- left_join(hybrid_df, jn_subjects, by = "issn") %>%
  select(doi, issn, top_level, subject_area)

# Number of Journals (full and fractional counting)
subject_area_all_jn <- subject_hybrid %>% 
  group_by(issn, subject_area) %>%
  summarise(n = n()) %>% 
  mutate(fractions = n / sum(n)) %>% 
  ungroup() %>% 
  group_by(subject_area) %>% 
  summarise(journals_n = n_distinct(issn),
            journals_frac = sum(fractions)) %>% 
  arrange(desc(journals_frac))
# Number of articles (full and fractional counting)
subject_area_all_article <- subject_hybrid %>% 
  group_by(doi, subject_area) %>%
  summarise(n = n()) %>% 
  mutate(fractions = n / sum(n)) %>% 
  ungroup() %>% 
  group_by(subject_area) %>% 
  summarise(articles_n = n_distinct(doi),
            articles_frac = sum(fractions)) %>% 
  arrange(desc(articles_frac))
```

Next, we present the hybrid OA uptake of different disciplines. Table
\ref{tab:subject_area_table} presents the high-level findings by ASJC
subject area. Between 2015 and 2019, the physical sciences
(`r filter(subject_area_all_jn, subject_area == "Physical Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Physical Sciences") %>% pull(articles_n)` OA articles) and health sciences  (`r filter(subject_area_all_jn, subject_area == "Health Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Health Sciences") %>% pull(articles_n)` OA articles) recorded the most
hybrid journals with at least one OA article. However,
the life sciences ((`r filter(subject_area_all_jn, subject_area == "Life Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Life Sciences") %>% pull(articles_n)` OA articles) published the most hybrid OA articles, while the
social sciences published the least (`r filter(subject_area_all_jn, subject_area == "Social Sciences") %>% pull(journals_n)`  hybrid journals, `r filter(subject_area_all_article, subject_area == "Social Sciences") %>% pull(articles_n)` articles). It is important to
note, though, that there is a large overlap between the journals' subject areas. Therefore, Table \ref{tab:subject_area_table} also presents fractional counts for hybrid journals and OA articles, which equally weights counting with respect to the total number of ASJC subject areas a journal belongs to. Journals with multiple ASJC codes that are members of the same subject area were counted once. For the remaining results, we only present the full counting.

```{r subject_area_table}
# join
inner_join(subject_area_all_jn, subject_area_all_article, by = "subject_area") %>%
  arrange(subject_area) %>%
  select(1, journals_n, articles_n, journals_frac, articles_frac) %>%
  filter(subject_area != "Multidisciplinary", !is.na(subject_area)) %>% 
  mutate(subject_area = stringr::str_to_sentence(subject_area)) %>%
  kbl(booktabs = TRUE, col.names = c("Subject area", "Journals", "OA articles", "Journals", "OA articles"),
      caption = "Full and Fractional Counting of Subject Areas (2015-2019).",
      format.args = list(big.mark = ",")) %>%
  add_header_above(c(" " = 1, "Full counting" = 2, "Fractional counting" = 2)) %>%
  kable_styling() 
```

```{r subject_ind}
# see 008_subject_exploration
hybrid_subject <- readr::read_csv(here::here("data", "jn_scopus_ind_subjects.csv"))
```

```{r}
subject_all_years <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(oa),
            n = sum(articles)) %>%
  ungroup() %>%
  group_by(description) %>%
  filter(subject_area != "Multidisciplinary")
# distinct journals per subject
journals_per_subject <- subject_all_years %>%
  summarise(n_journals = n_distinct(issn)) %>%
  mutate(field_label = glue::glue("{description} (n={n_journals})"))

median_all <- subject_all_years %>% 
  ungroup() %>% 
  distinct(issn, .keep_all = TRUE) %>% 
  mutate(prop = n_oa /n) %>% 
  .$prop %>% 
  median()
# rank
oa_per_subject_rank <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(oa),
            n = sum(articles)) %>%
  mutate(prop = n_oa / n) %>%
  ungroup() %>% 
  group_by(description) %>%
  summarise(median_prop = median(prop)) %>%
  arrange(desc(median_prop))
```

Figure \ref{fig:oa_sub_uptake} presents the journals' five-year OA uptake grouped by subject area and field and in comparison to the overall median (Mdn= `r median_all * 100`; excluding one journal coded as multidisciplinary). The box plot reveals some variation between the subject areas. Notably, with the exception of environmental science and earth and planetary sciences, journals from the physical sciences show a lower median proportion of hybrid OA articles, whereas most life sciences and social sciences journals ranked above average. Figure \ref{fig:oa_sub_uptake}  also shows large variations among the 26 subject fields, with the highest OA rates in immunology and microbiology (Mdn=`r filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop)*100`), environmental science (Mdn=`r filter(oa_per_subject_rank, description == "Environmental Science") %>% pull(median_prop)*100`), and neuroscience (Mdn=`r filter(oa_per_subject_rank, description == "Neuroscience") %>% pull(median_prop)*100`), whilst  chemistry (Mdn=`r filter(oa_per_subject_rank, description == "Chemistry") %>% pull(median_prop)*100`), materials science (Mdn=`r filter(oa_per_subject_rank, description == "Materials Science") %>% pull(median_prop)*100`), and dentistry (Mdn=`r filter(oa_per_subject_rank, description == "Dentistry") %>% pull(median_prop)*100`) recorded the lowest OA rate.

```{r oa_sub_uptake, fig.cap='OA Uptake in Elsevier Hybrid Journals by Subject Area and Field (2015-2019).', fig.asp=1.3, out.width="90%"}
# with labels
subject_all_years_ <- 
  inner_join(subject_all_years, journals_per_subject, by = "description")

p_subject <- ggplot(subject_all_years_, aes(fct_reorder(field_label, n_oa / n, .fun = median), n_oa / n)) + 
  geom_boxplot(aes(color = subject_area)) +
  geom_jitter(alpha = .05, aes(color = subject_area), size = 1) +
  facet_wrap(~subject_area, scales = "free_y", nrow = 4) +
  geom_hline(yintercept = median_all,  colour = "red", linetype ="dashed", size = 1) +
  scale_colour_manual(NULL, values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701"),
                      guide=FALSE) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 5L),
    expand = expansion(mult = c(0, 0.05))
  ) +
  coord_flip(ylim=c(0,0.2)) +
  labs(x = NULL, y = "Five-Year Hybrid OA Uptake Rate\nper Journal") +
  theme_minimal_vgrid(font_size = 10) +
  theme(strip.text = element_text(face="bold"))
p_subject
```

## Invoicing for Hybrid Open Access{.unnumbered}

### Invoice Channels{.unnumbered}

```{r}
# load eligible oa articles
hybrid_volume <- readr::read_csv(here::here("data", "hybrid_articles.csv")) 
# add open apc info
hybrid_volume <- hybrid_volume %>%
      mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
# aggregate data
sponsored_articles <- hybrid_volume %>%
  mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND")) %>%
  count(issued_year, fct_source, cc_norm)
all_articles <- hybrid_volume %>%
  group_by(issued_year) %>%
  summarise(n = n())
sponsored_art <- sponsored_articles %>% 
  group_by(fct_source) %>% 
  summarise(nn = sum(n)) %>% 
  #mutate(cumsum = cumsum(nn)) %>% 
  mutate(prop = nn / sum(nn) * 100)  %>%
 # mutate(cumprop = cumsum(prop)) %>%
  adorn_totals("row")
```

As can be seen from Table \ref{tab:invoicing_overview}, hybrid OA APCs were most often invoiced to authors (n=`r filter(sponsored_art, fct_source == "Author") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Author") %>% pull(prop)`%) and to a lesser extent as part of agreements (n=`r filter(sponsored_art, fct_source == "Agreement") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Agreement") %>% pull(prop)`%). Interestingly, we also found a small number of cases where hybrid APCs were waived (n=`r filter(sponsored_art, fct_source == "Fee waived") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Fee waived") %>% pull(prop)`%). A brief note of caution regarding APCs invoiced to the authors is due here: Although author invoicing was indicated in the \texttt{openaccessSponsorType} article metadata field used for this study, this does not necessarily mean that authors paid the APC out of their own pocket. Rather, this implies that the authors' institution or funder did not have a central invoicing agreement in place at the time of publication.

```{r invoicing_overview}
sponsored_art %>% 
  kbl(booktabs = TRUE, col.names = c("Invoicing channel", "Hybrid OA articles (n)", "Percentage"),
      caption = "Invoice Channels for Elsevier Hybrid OA Articles (2015-2019).", digits = 1,
      format.args = list(big.mark = ",")) %>%
  kable_styling() %>%
  row_spec(4, hline_after = TRUE)
```

```{r els_waived}
els_waived_df <- hybrid_volume %>% 
  filter(oa_sponsor_type == "ElsevierWaived") 
sponsored_articles_prop <- sponsored_articles %>%
  group_by(issued_year, fct_source) %>%
  summarise(n = sum(n)) %>%
  mutate(perc = n / sum(n) * 100)
```

Figure \ref{fig:invoiceoverview} illustrates that over the years, Elsevier has consistently invoiced more authors than funders or academic consortia ("Agreement"), although both shares increased. In particular, the proportion of author-invoicing increased from `r filter(sponsored_articles_prop, issued_year == 2015, fct_source == "Author") %>% pull(perc)`% to `r filter(sponsored_articles_prop, issued_year == 2019, fct_source == "Author") %>% pull(perc)`%, whereas the share of invoicing through agreements slightly
increased from `r filter(sponsored_articles_prop, issued_year == 2015, fct_source == "Agreement") %>% pull(perc)`% to `r filter(sponsored_articles_prop, issued_year == 2019, fct_source == "Agreement") %>% pull(perc)`%, respectively. The share of fee-waived articles also remained relatively stable, but we found different types of waivers. Around `r filter(els_waived_df, !is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)* 100`% of fee waivers were linked to an OA sponsor, which indicates that a third party paid the APC. For instance, `r filter(els_waived_df, oa_sponsor_name == "Académie des Sciences") %>% distinct(doi) %>% nrow()` waivers were linked to the French Académie des Sciences, presumably covering OA publication in its society journals for affiliated authors. The remaining `r filter(els_waived_df, is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)*100`% of waived articles did not disclose any OA sponsor. Moreover, Figure \ref{fig:invoiceoverview} compares the invoicing channels based on open content license variants. When Elsevier invoiced authors directly, most OA articles in hybrid journals were published under a non-commercial license (n=`r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC BY-NC-ND") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC BY-NC-ND") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Author") %>% pull(n) %>% sum()* 100`%), whereas most articles billed as part of agreements were licensed under the more permissive CC BY license (n=`r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC BY") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC BY") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Agreement") %>% pull(n) %>% sum() * 100`%).

```{r invoiceoverview, fig.cap="Elsevier Hybrid OA Articles by Invoice Channel and CC License (Per Year)."}
# plot
sponsored_articles %>%
  mutate(fct_source = stringr::str_to_title(fct_source)) %>%
  mutate(fct_source = ordered(fct_source, levels = c("Author", "Agreement", "Fee Waived", "Other"))) %>%
  ggplot(aes(x = gsub("20", "'", issued_year), y = n), alpha = 0.8) +
  geom_bar(
    data = all_articles,
    color = "transparent",
    stat = "identity",
    fill = "#b3b3b3a0"
  ) +
  geom_bar(aes(fill = cc_norm), color = "white", stat = "identity", alpha = 0.8, size = 0.25,
           position = position_stack(reverse = T)) +
  facet_wrap( ~ fct_source, nrow = 1) +
  scale_fill_manual(values = 
                      c(`CC BY` = "#B52141",
                        `CC BY-NC-ND` = "#0093c7"),
                    name = "CC License Variant") +
  labs(x = "Publication Year", y = "Hybrid OA Articles") + 
  scale_y_continuous( 
                     labels = scales::number_format(big.mark = ","),
                      expand = expansion(mult = c(0, 0.05))) +
  theme_minimal_hgrid() +
  theme(legend.position="top", legend.justification = "right")
```

We also observed large differences in OA invoicing among subject fields.
Table \ref{tab:invocing_by_subject} shows the number of OA articles by
subject field and invoice channel. For articles in nursing, decision
sciences, and pharmacology, toxicology and pharmaceutics, Elsevier
predominantly invoiced authors, whereas most energy and chemical
engineering articles were invoiced through agreements. Likewise, the
majority of articles in materials science, chemistry and physics and
astronomy were not invoiced to authors but facilitated through
agreements or waived. The large share of waived APCs in physics and
astronomy can be attributed to a single 2015 issue of Nuclear and
Particle Physics Proceedings.

```{r invocing_by_subject, results='asis'}
hybrid_jn_volume <- hybrid_oa_volume %>%
  group_by(issn) %>% 
  summarise(n_oa = sum(oa))

jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))

hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
funder_shares <- hybrid_df %>%
   mutate(oa_sponsor_type = ifelse(!oa_sponsor_type %in% c("FundingBody", "Author", "ElsevierWaived"), "Other", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  group_by(oa_sponsor_type, issn) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  left_join(hybrid_jn_volume, by = "issn") %>%
  mutate(prop = n_oa_type / n_oa) %>%
  left_join(jn_subjects, by = "issn") %>%
  ungroup() %>%
  group_by(top_level, oa_sponsor_type) %>%
  summarise(oa_type = sum(n_oa_type)) %>%
  mutate(prop = oa_type / sum(oa_type))
# overall props
oa_sponsor_type_all <- hybrid_df %>% 
  #exclude multidisciplinary
  filter(issn != "2095-9273") %>%
  count(oa_sponsor_type) %>% 
  mutate(prop = n / sum(n))
subject_all <- funder_shares %>%
  group_by(top_level) %>%
  summarise(oa = sum(oa_type))

funder_overview_subject <- left_join(funder_shares, subject_all, by = "top_level") %>%
  tidyr::pivot_wider(id_cols = c("top_level", "oa"), names_from = oa_sponsor_type, values_from = prop) %>%
  ungroup() %>% 
  # remove na and multidisciplinary 
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  relocate(Agreement, .after = Author) %>%
  arrange(desc(Author))

# define heat-colors
make_pal <- scales::col_numeric(
  palette  = c("white", "#fcde9c","#faa476","#f0746e","#e34f6f","#dc3977"),
  na.color = "#808080",
  domain = NULL,
  alpha = FALSE,
  reverse = FALSE
)
# prepare row-wise grouping per subject area
subject_table <- jn_subjects %>% 
  distinct(subject_area, top_level) %>%
  inner_join(funder_overview_subject, by = "top_level") %>%
  mutate_at(4:7, function(x){
    round(x*100,0)
  }) %>%
  mutate_at(4:7, function(x){
    cell_spec(x, background = make_pal(x))}) %>%
  mutate_at(3, function(x){
    cell_spec(x, background = make_pal(x))})

subject_table_ <- subject_table %>%
  group_by(subject_area) %>%
  arrange(top_level, .by_group = TRUE) %>% 
  ungroup() 
subject_table_ %>%
  select(-subject_area) %>%
            kbl(booktabs = TRUE, "latex", format.args = list(big.mark = ","), digits = 1, escape = FALSE,
                  caption = "Invoice channels by discipline 2015-2019.",
                  position = "H",
      col.names = c("Subject", "OA articles", "Author", "Agreement", "Fee waived", "Other"),
       linesep = "",
      align = "lrrrrr") %>%
  add_header_above(c(" " = 2, "Invoice channel (in %)" = 4)) %>%
  pack_rows(index = table(subject_table_$subject_area)) %>%
  kable_styling(latex_options = c("scale_down"))
```

### Invoice Recipients{.unnumbered}

```{r}
hybrid_invoice <- hybrid_volume %>%
  filter(oa_sponsor_type == "FundingBody") %>%
  left_join(curated_invoicing_df, by = c("oa_sponsor_name")) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND"))
```

```{r license_count}
sponsor_df <- hybrid_invoice %>% 
#  filter(!is.na(Country)) %>% 
  # manual changes
  mutate(Country = ifelse(grepl("EU", Country), "European Union", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf))
```

Elsevier’s data offer more insight into invoicing. Overall, we identified `r filter(sponsor_df, !is.na(oa_sponsor_cleaned)) %>% distinct(oa_sponsor_cleaned) %>% nrow()` academic institutions and funders that received
invoices as part of publishing agreements. By a large margin, most invoices were issued to
UK-based research funders and institutions (n=`r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%), followed by the Netherlands (n=`r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%) and the European Union (n=`r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%; EU), which in the context of this paper refers to institutions established by the European Commission, and not the  individual member states. Invoicing data comprises two EU institutions, the research funder European Research Council and the health agency European Center for Disease Prevention and Control (ECDS). Figure \ref{fig:invoice_sponsor_country} presents the geographical distribution by license prevalence,
highlighting the dominance of CC BY licensed articles invoiced to institutions in the UK, the United States (US), and, to a much
lesser extent, the Netherlands. In contrast, most articles
invoiced to institutions from Norway or representing the EU
were published under a non-commercial license.

```{r invoice_sponsor_country, fig.cap="Centrally Invoiced Elsevier Hybrid OA Articles by Country and CC License (2015-2019)."}
sponsor_df %>%
  group_by(country_factor, cc_norm) %>% 
  summarise(n = n_distinct(doi)) %>%
  ggplot(aes(fct_rev(country_factor), n, fill = cc_norm)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE),
            color = "white", alpha = 0.8, size = 0.25) +
  coord_flip() +
  scale_fill_manual(values = 
                      c(`CC BY` = "#B52141",
                        `CC BY-NC-ND` = "#0093c7"),
                    name = "CC License Variant") +
  labs(x = "OA Sponsor by Country", y = "Hybrid OA Articles (Agreement)") +
  scale_y_continuous(labels = scales::number_format(big.mark = ","),
                      expand = expansion(mult = c(0, 0.05)),
                     breaks = seq(0,15000, 2000)) +
  theme_minimal_vgrid() +
  theme(legend.position="top", legend.justification = "right")
```

Figure \ref{fig:sponsor_sector} shows the yearly distribution of centrally
invoiced articles by year and institutional sector. While UK-based OA
invoices were mainly addressed to discipline-specific governmental and
non-profit research funders, invoices to the Netherlands and Norway were
issued to national academic consortia representing the higher education
sector. In 2019, Elsevier also launched similar agreements in countries
with lower publication output including Hungary and Poland. We
found that invoice recipients from the UK and the US mainly
represented discipline-specific funders, while invoice recipients from other countries focused on a broad variety of
disciplines.

```{r sponsor_sector, fig.cap='Centrally Invoiced Elsevier Hybrid OA Articles by Country and Sector (Per Year).'}
sponsor_df %>%
  filter(!is.na(Sector)) %>%
  count(issued_year, Sector, country_factor) %>%
  mutate(Sector = stringr::str_to_title(Sector)) %>% 
  mutate(sector_level = factor(Sector)) %>%
  mutate(sector_level = fct_relevel(sector_level, "Government", "Higher Education", "Private Non-Profit", "Other")) %>%
  ggplot(aes(gsub("20", "'", issued_year), n, fill = sector_level)) + 
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE),
            color = "white", alpha = 0.8, size = 0.25) +
  facet_wrap(~country_factor) +
  scale_fill_manual("Sector", values = c("#65BADA", "#00A757", "#E5BA52", "grey50")) +
  scale_y_continuous(labels = scales::number_format(big.mark = " "),
                      expand = expansion(mult = c(0, 0.05))) +
  theme_minimal_hgrid() +
  theme(legend.position="top", legend.justification = "right",
        panel.spacing = unit(2, "lines")) +
  guides(fill = guide_legend(nrow = 2)) +
  labs(x = "Publication Year", y = "Hybrid OA Articles (Agreement)")
```

```{r}
sponsor_country <- hybrid_invoice %>%
  filter(!is.na(`Primary research focus`)) %>%
  # manual changes
  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "European Union", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf)) %>%
  mutate(`Primary research focus` = ifelse(`Primary research focus` == "SSH", "SSH", stringr::str_to_sentence(`Primary research focus`))) %>%
  group_by(country_factor, `Primary research focus`) %>%
  summarise(n = n_distinct(doi)) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n), 2) * 100)
```

```{r open_apc_check}
hybrid_invoice <- hybrid_invoice %>%
  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
```

Table \ref{tab:sponsor_top_10} presents the top ten of `r hybrid_invoice %>% filter(!is.na(oa_sponsor_cleaned)) %>% distinct(oa_sponsor_cleaned) %>% nrow()` sponsoring bodies in our sample. Together, they accounted for around 80\% of
centrally invoiced APCs. The table also highlights the number of
distinct journals and the share of CC BY licensed articles. Notably,
UK-based research funders and the US-based Bill and Melinda Gates
Foundation mainly received invoices for CC BY licensed articles. In
contrast, articles invoiced to VSNU, the European Research Council, and
Norway Institutes had a much lower proportion of CC BY.

```{r, sponsor_top_10}
hybrid_invoice_tt <- hybrid_invoice %>%
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_cleaned), "Other", oa_sponsor_cleaned)) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND")) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
hybrid_invoice_agg_general <- hybrid_invoice_tt %>%
  group_by(oa_sponsor) %>%
  summarise(journals = n_distinct(issn),
            articles = n_distinct(doi)) %>%
  mutate(prop = articles / sum(articles))
hybrid_invoice_agg_cc <- hybrid_invoice_tt %>%
  ungroup() %>%
  filter(cc_norm == "CC BY") %>%
  group_by(oa_sponsor) %>%
  summarise(cc_by = n_distinct(doi))
hybrid_invoice_agg_oapc <- hybrid_invoice_tt %>%
  filter(in_open_apc == TRUE) %>%
  group_by(oa_sponsor) %>%
  summarise(in_open_apc =  n_distinct(doi))
ind_sponsor_df <-
  left_join(hybrid_invoice_agg_general, hybrid_invoice_agg_cc, by = "oa_sponsor") %>%
  left_join(hybrid_invoice_agg_oapc, by = "oa_sponsor") %>%
  mutate(
    cc_by = cc_by / articles,
    in_open_apc = ifelse(is.na(in_open_apc), 0, in_open_apc),
    in_open_apc = in_open_apc / articles
  ) %>%
  mutate(oa_sponsor = fct_reorder(oa_sponsor, articles)) %>%
  mutate(oa_sponsor = forcats::fct_relevel(oa_sponsor, "Other", after = 0L)) %>% arrange(desc(oa_sponsor))
# overall
overall_ind_sponsor_df <-
  tibble(
    oa_sponsor = "All",
    journals = length(unique(hybrid_invoice$issn)),
    articles = length(unique(hybrid_invoice$doi)),
    prop = 1,
    cc_by = filter(hybrid_invoice, cc_norm == "CC BY") %>% nrow() / length(unique(hybrid_invoice$doi)),
    in_open_apc = filter(hybrid_invoice, in_open_apc == TRUE) %>% nrow() / length(unique(hybrid_invoice$doi))
  )

ind_sponsor_df %>%
  bind_rows(overall_ind_sponsor_df) %>%
  mutate(oa_sponsor = gsub("&", "and", oa_sponsor)) %>%
  mutate_at(4:6, function(x){
    round(x*100,0)
  }) %>%
  mutate_at(5:6, function(x){
    cell_spec(x, background = make_pal(x))}) %>%
  kbl("latex", booktabs = TRUE, escape = FALSE,
      col.names = c("Hybrid OA sponsors", "Journals", "Articles", "\\%", "CC BY", "OAPC"),
      align = c("lrrrrr"),
      linesep = "",
      caption = "Elsevier Hybrid OA Invoice Recipients  (2015-2019).") %>%
  row_spec(11:12, hline_after = TRUE) %>%
  add_header_above(c(" " = 4, "Compliance (in %)" = 2)) %>%
   kable_styling(latex_options = c("scale_down"))
```

The table also highlights the proportion of APC funding publicly disclosed through the Open APC initiative, showing higher disclosure rates for  funders with a large CC BY share. On the other hand, APCs that were invoiced as part of transformative agreements were not publicly shared.

# Discussion

From 2015-2019, Elsevier recorded growth in the uptake of
hybrid OA: The number of hybrid OA
articles published per year doubled, the number of hybrid journals with
at least one OA article grew by 21\%, and the share of hybrid OA
articles relative to closed-accessed articles in these journals
increased from 2.6\% to 3.7\%. As  @Laakso_2016, we observed disciplinary differences. In particular, we found the highest count of hybrid OA
articles in physical sciences journals (see Table
\ref{tab:subject_area_table}). This was followed by the life sciences
and health sciences, whereas the social sciences had the lowest count. This order mostly reflects the disciplines' overall
publication output. According to the Open Science Monitor [@OS_Monitor], the physical sciences publish the
most articles, followed by the health sciences, life sciences, and
social sciences.

Disciplinary differences in hybrid OA prevalence become more
meaningful when considering the relative share of OA articles to
closed-access articles. In line with previous research, we found that
Elsevier journals from the life sciences and social sciences  [@Jubb_2017; @Kramer_2018; @Laakso_2016] recorded
greater than typical hybrid OA uptake, whereas physical sciences
journals generally had a lower than typical uptake [@Kramer_2018; @Mart_n_Mart_n_2018; @Laakso_2016]. In a
systematic review of disciplinary OA publishing patterns, @Severin_2020
highlighted the importance of socio-cultural and technological
factors in shaping publishing cultures and practices.  For instance, since many
branches of the physical sciences have established self-archiving
practices [@Severin_2020, @Bj_rk_2010, @Laakso_2016], researchers can provide OA through repositories and might perceive
much less of a need for hybrid OA. In contrast, hybrid OA in the life sciences could be enabled through project-based funding structures as they allow for easy integration of publishing costs @Severin_2020.
On the other hand, the high hybrid OA uptake that we observed among the social sciences could point to the influential role of OA policies and invoicing agreements [@Huang_2020; @Larivi_re_2018]. While these aspects
are important drivers of both full and hybrid OA, faculty members
generally prioritize journal prestige and the journal impact factor
 [@Niles_2020] in publishing decisions, which, in combination with
the availability of OA funding, might encourage publishing hybrid OA
[@van_der_graaf_maurits_2017].

Our invoicing data analysis found that most APCs were
invoiced to the author (n=41,725; 58.2\%). However, it is important to
emphasize although Elsevier's metadata classifies the sponsor type as
``author'', this does not necessarily mean that authors paid for APCs themselves but rather that APCs were not invoiced
through publishing agreements. It is possible that these APCs were covered through institutional funds or  research
grants. This would align with a recent Springer Nature survey that found hybrid
OA was predominantly supported through institutional and funder sources
(71\%), followed by OA agreements (34\%), while only 6\% were paid from
personal funds or savings [@Monaghan_2020].

Further, we observed
notable differences in licensing. Most
hybrid OA articles invoiced to author were licensed under
the more restrictive CC BY-NC-ND license. Previous research, while lacking dedicated
studies on license selection,  suggests that authors tend to select more restrictive license variants when
given a choice [@Fraser_2020; @Rowley_2017; @Van_Noorden_2013]. On the other hand, we found when hybrid OA was invoiced
through agreements, most articles were licensed under the more liberal
CC BY license. As several funding bodies mandate CC BY licenses,
including the UK research funders that account for 62\% of our
"agreement" subsample, this result is perhaps not surprising but
suggests the effectiveness of such agreements.

Based on our findings, around a third of the articles were invoiced
 through OA agreements (e.g., research funders, national library
consortia). The predominance of UK funding bodies in this subsample reiterates reports from previous
studies that the UK's OA profile differs markedly from other countries
(cf. @Jubb_2015; @Jubb_2017), pointing to the impact of science policy. To promote 
OA, the UK implemented centralized APC funding
and embraced hybrid OA as a transition model. Within five years of the
publication of the Finch Report in 2012, the UK recorded an 18\%
increase in immediate OA, coupled with a rise in hybrid OA from 2.7\% in
2012 to 15.4\% of all articles in 2016 [@Jubb_2017], which can be
attributed to the availability of funding from 
RCUK. Indeed, RCUK block grants have been the largest single source of
APC funds in the UK---the Wellcome Trust, and more recently
transformative agreements [@Tickel_2018; @Jubb_2017]. However,
since then publishing expenditures of UK universities have been rising
rapidly (Jubb et al., 2017), so that several universities stopped supporting hybrid OA
in late 2018-2019 [@walker_2019; @birmingham]. This development might
explain the slight decrease in hybrid articles invoiced to the UK our
data showed around that time (see Figure
\ref{fig:sponsor_sector}).

Through this study we
demonstrated the utility and benefits of publisher-provided
metadata about hybrid OA invoicing and highlighted the need for extending
these to include comprehensive information about licensing and APC waivers. Metadata guidance should also consider the substantial amount of delayed OA content, which needs to be distinguished from hybrid OA.
Hence, our study substantiates the recommendations from the ESAC Initiative that seek to increase efficiency and
transparency through improved invoicing and
reporting processes and metadata about OA funding [@Geschuhn_2017; @Marques_2019].

While this study advances our knowledge about hybrid OA uptake and
invoicing, the limitations leave room for future research. We
focused on only one publisher, which limits the generalizability of our
findings because Elsevier's journal
portfolio, mix of business models, pricing, and promotion of various
options cannot be assumed to be representative of scholarly journal
publishing in general. Relatedly, our study is limited to articles published in hybrid OA journals and, hence, is not representative of invoices for articles in full OA journals. To develop a full picture of OA invoicing, the methodology presented in this paper can be used for future studies in this direction. Further, although Elsevier's invoicing data improves transparency, it seems likely that not all actual OA funding bodies are disclosed. For instance, most articles were invoiced to authors, but it remains unclear
if the APCs were paid by the authors themselves or
through institutional OA funds or research grants. Research into
this topic would improve our understanding of OA funding 
outside of publishing agreements. Moreover, our study demonstrates that
comprehensive mapping of the financial
flows of OA publishing requires complex, in-depth country-specific
analyses. Such studies would ideally draw on
various data sources to consider  research funders, the 
consortia landscape, OA policies, and  publishing agreements Future research could analyze funder policies in combination with institutional agreements (including which journals are covered by the agreements) and hybrid OA invoicing metadata to provide more insight into publishing patterns and CC licence choice. A further avenue of research would be dedicated studies on the strengths and weaknesses of data provided by publishers and the Open APC initiative, and consider an integrative approach that combines both datastreams to reap the benefits of each while mitigating their individual drawbacks.

The policy implications and the underlying data of this study present opportunities for actors in the scholarly publishing landscape to rely less on assumptions and secondary data, and instead establish a direct connection between the publication output and the related financial information. This could increase accountability in spending public and charitable funds, and improve OA cost monitoring through enhanced transparency, precision, and timeliness. Analyses like we have presented here can inform library consortia in negotiations through better understanding their publisher-specific OA output, determining the amount of delayed OA content, and facilitating comparative analyses across different consortia. The study could also inform research funders and research institutions about license selection implications if there is not any strict policy in place. This study provides a snapshot of hybrid OA for Elsevier, the largest journal
publisher, prior to the impact of the implementation of Plan S, an initiative to accelerate the transition to OA that will no longer support hybrid OA [@Plan_s]. While many Plan S signatories have already had strong OA
policies, this harmonized approach is likely to affect publishing
decisions of funded authors, the licencing of their articles, and the
offerings and pricing of publishers at a larger scale than before. Because the new requirements apply to research funded from 2021 onward, a comparative study
on articles invoiced to Plan S signatories would be a fruitful
endeavor. Plan S does not fund hybrid OA publication unless the host journal is explicitly committed to convert to full OA, however, with the growth of national and institutional agreements that enable affiliated authors to publish hybrid OA, such restrictions will influence researchers differently. Invoicing data as used for this study can be useful in exploring later on, by looking at research funding statements and invoicing channels in relation to each other.

# Conclusion

The primary aim of this empirical study was to investigate Elsevier's
hybrid OA publishing from 2015-2019 to better understand the volume and
invoicing of hybrid OA and to present a novel, data-driven approach for
such analyses. Our results indicate that although the number of hybrid
OA articles has increased over time, its uptake has remained low.
Notably, hybrid APCs were most often invoiced directly to the authors, followed by agreements, where only a few funding bodies were
the primary drivers of hybrid OA. Finally, our findings highlight that
publisher-provided metadata about the invoicing channels of (hybrid) OA
can facilitate research into and increase the understanding of the
financial flows of OA publishing.

Since the beginning, hybrid OA has been a challenging subject to
study due to the lack of standardized ways publishers flag such
content and APC funding data being limited self-reported data, surveys,
and other secondary sources. This study presented a novel
approach to studying APC invoicing that is based on publicly
available publisher-provided metadata, which can be used on on its own
or in combination with other public data sources to gain more detailed
and comprehensive insights into hybrid OA uptake and invoicing. If more publishers reported OA invoicing on the article
level and in a machine-readable format, this would increase transparency
and improve monitoring of the scholarly journal landscape over
time. As hybrid OA has  become a central element of OA
policies of research funders and libraries, consumer organizations could
require that invoicing information is added to the article-level
metadata. As long as publishers do not provide this data in a structured
and comprehensive format, they prevent benchmarking prices and therefore hinder competition.

From recent science policy developments in Europe it appears that Big
Deals have gained support and remain firmly in place in the form of
transformative agreements. Moreover, such agreements seem to become an increasingly dominant feature of scholarly publishing as indicated by the number of contracts that Elsevier and other publishers have successfully negotiated and secured since we conducted the analyses presented here.[^1] Through this study we can affirm that hybrid OA is complex
as the financial flows involve research funders, libraries, consortia, and authors. However, it is on
 publishers to increase the transparency of OA publishing, including hybrid OA and transformative agreements, by providing instantaneous open data about OA uptake and invoicing.

# Data Availability

The source code data analysis is available on GitHub:

<https://github.com/njahn82/elsevier_hybrid_invoicing>

# References

[^1]: See, for instance, the announcements from the University of California: <https://www.universityofcalifornia.edu/news/uc-s-deal-elsevier-what-it-took-what-it-means-why-it-matters> and the Danish National Licensing Consortium: <https://www.kb.dk/en/about-us/licensing/national-licensing-consortium/elsevier-agreement-2021-24>